### 2016-08-05 paper.tex
updating with the KBS format version

%\documentclass[final,1p,times]{elsarticle}
%\documentclass[twocolumn]{elsarticle}
\documentclass[a4paper,10pt,twocolumn,preprint,3p]{elsarticle}

\usepackage{graphics}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[dvips]{epsfig}
\usepackage[latin1]{inputenc}
\usepackage{url}
\usepackage{appendix}

%\def\CC{{C\hspace{-.05em}\raisebox{.4ex}{\tiny\bf ++}}~}
%\addtolength{\textfloatsep}{-0.5cm}
%\addtolength{\intextsep}{-0.5cm}

\journal{Knowledge-Based Systems}

\begin{document}

\begin{frontmatter}

%%%%%%%%%%%%%%%% Titulo %%%%%%%%%%%%%%%
\title{Applying Computational Intelligence Methods for Predicting the Sales of Newly Published Books in a Real Editorial Business Management Environment}

%\setlength\textwidth\columnwidth

%%%%%%%%%%%%%%%% autores %%%%%%%%%%%%%%
\author[ugr]{Pedro A. Castillo}
\ead{pacv@ugr.es}
\author[ugr]{Antonio M. Mora}
\ead{amorag@geneura.ugr.es}
\author[abd]{H. Faris}
\ead{hossam.faris@ju.edu.jo}
\author[ugr]{J.J. Merelo}
\ead{jmerelo@geneura.ugr.es}
\author[ugr]{Pablo Garc\'{\i}a-S\'anchez}
\ead{pablogarcia@ugr.es}
\author[ugr]{Antonio J. Fern\'andez-Ares}
\ead{antares.es@gmail.com}
\author[ugr]{Paloma De las Cuevas}
\ead{palomacd@ugr.es}
\author[ugr]{Mar\'ia I. Garc\'ia-Arenas}
\ead{mgarenas@ugr.es}

\address[ugr]{Department of Computer Architecture and Computer Technology, ETSIIT and CITIC \\
University of Granada, Granada, Spain. Tel: +34958241778. Fax: +34958248993}
\address[abd]{Business Information Technology Department, King Abdullah II School for Information Technology \\
The University of Jordan, Amman, Jordan}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
When a new book is launched the publisher faces the problem of how
many books should be printed for delivery to bookstores; printing too
many is the main issue, since it implies a loss of investment due to
inventory excess, but printing too few will also have a negative economic impact. 
In this paper, we are tackling the problem of predicting total sales 
in order to print the right amount of books and doing so even before the book has reached the stores. A real dataset including the complete sales data for books published in Spain across several years has been used.
We have conducted an analysis in three stages: an initial exploratory analysis, by means of data visualisation techniques; a feature selection process, using different techniques to find out what are the variables that have more impact on sales; and a regression or prediction stage, in which a set of machine learning methods has been applied to create forecasting models for book sales. 
The obtained models are able to predict sales from pre-publication data with
remarkable accuracy, and can be visualised as simple decision trees. 
Thus, these can be used as decision-aid tools for publishers, which can provide 
a reliable guidance on the decision process of publishing a book. 
This is also shown in the paper by addressing four example cases of representative 
publishers, regarding their number of sales and the number of different books they sell.
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
Book sales forecasting \sep Decision-aid models \sep Feature selection \sep Regression  
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\end{keyword}

\end{frontmatter}

%********************************************************************************

\section{Introduction}

Publishing a book, in the same way as releasing any other cultural product with a physical substrate, implies several types of risks and costs due to
the complexity, and derived expenses, of the production,
distribution, and storage processes; most of these costs are associated
with the number of books actually printed. But the problem is further
complicated for completely new products, in the sense that there is an unknown
author or a new topic with no track record. Sales forecasts can
only be based on the knowledge obtained from other similar products,
but the definition of ``similar'' itself is fuzzy and, sometimes,
subjective. Even if the author or the topic is known, doing accurate
estimations of future sales of the new product is quite difficult, and extremely important, so that production runs heed predicted sales. 
Indeed, several studies have demonstrated that improving sales predictions 
(a reduction of forecasting errors) results in an enhancement in the 
production process \cite{Fildes2010,Saeed2008}.

This problem is specially acute for book publishers, who, among other
content providers, release new {\em products} more frequently than
other industries, so that they face the problem of predicting sales quite often. The issue, in this case, is to print an adequate number of copies but not too many, as the unsold volumes will lead to losses and sunk inventory cost, 
whereas if the number of printed copies was not enough, a new print
run can be made, resulting in temporary losses due to a lack of supply or in bad marketing for the customers, for instance.

Despite the fact that errors in one or the other direction do not have the same impact, it is extremely important to develop an accurate predictive method
which could forecast the future sales a new book will achieve, in order to optimise production schedules, improve the publishing company profits and minimise losses \cite{Zhao2001}, being the latter the main objective of this paper.

It is important to notice that the accuracy of the predictive method that we intend to design is affected negatively by several factors related to decisions on purchases, to freight traffic, or with the distribution strategy \cite{Little1998}.
Thus, it is mandatory to know the product and the variables that
might affect the sales, and how an expert manager takes them into 
account \cite{Armstrong2001,SThomassey2014}. 
Moreover, their impact is difficult to estimate and it is not constant over time, hence the complexity to adequately identify and measure their influence \cite{DeToni2000}.
In addition, some variables may have correlations between them, although these correlations can be detected using data analysis methods, 
such as Self-Organizing Maps (SOM) \cite{kohonen1998} or stepwise regression analysis (SRA) \cite{Chang2009}.

In Spain, the process of releasing a book goes like this: the
publisher decides, based on past experience, how many books to
print. These books are distributed to points of sale, but also given
out to reviewers and literary magazines, the number of these will
depend on how many you have printed. Books then go, in some cases, to
the {\em novelty} table, that is, prominently displayed in the
bookstore, shown in the window, or highlighted using props or other
kind of advertisement. Books stay for sale for as long as the
publisher or the store want until the bookseller decides to return it
to the publisher or just, in some cases, move it away from shelves to
storage. Books returned to the publisher are pulped, sold back or
given to the author or distributed in second-hand channels such as
book stalls or second-hand bookstores. If the print run has been fully
distributed and there are more requests from bookstores, on the other
hand, a new print run is made and distributed to bookstores. 
Even if the process has changed
in the last few years since Amazon started to sell in Spain, which
actually happened {\em after} the data used for this paper was
collected, it roughly stays the same. Other countries will have a
similar model, although scale will be different.

Our intention in this paper is to find out the main factor influencing
sales in order to give a tool that the publisher can use to decide how
many books should be printed, as well as how to leverage these printed
copies to maximize sales using the decision variables under his
control. That is why, using data obtained from a  company that sells
software for publishers, we analyse it by means of different data
mining methods and create predictive models that can be mainly used as
decision-aid tools for book publishers. With these models, publishers
will be able to 
combine their expert knowledge about the market with the created
forecasting models in order to get a reliable estimation of book
sales, and, based on it, act upon it to maximize the books sold for a
particular print run.

In the research published in this paper, the problem 
of predicting total sales in order to print the right amount of books
%of how many units should be printed  when a new book is published 
is faced using real data provided by the 
Spanish publishing company Trevenque Editorial S.L.\footnote{{\tt http://editorial.trevenque.es}}. It provides management systems and services for bookshops, publishers, and distributors. 
Usually, the company's expert sales managers analysed historical data and applied their knowledge using external variables to compute correct estimations (predictions); for instance, they can decide increasing or decreasing the sales forecast based on season or sales period.
Nevertheless, just carrying out the analysis and forecasting process as a human expert may lead to imprecise predictions, and they may not take all the variables into account. 
In addition, the process might be tedious, especially when the amount
of data is quite large. Moreover, different experts can reach different
predictions from the same dataset \cite{Sanders1994}.

%An initial version of this method was developed under a contract with that company and eventually integrated in their product. This paper represents an improved version of that method based on the same data. 
% We should also mention what data we actually mean, I don't think
% we've published it here. - JJ
% Antonio - I think this information is not relevant for the paper...

With this issue in mind, in this paper we firstly present an analysis on the aforementioned real dataset, by means of data mining and visualisation techniques. Then, a feature selection process is conducted using three different methods, in order to find out what are the relevant variables describing a book in the prediction (or regression) procedure. After this, a forecasting stage is performed applying a set of regression algorithms to the different datasets, also considering a different number of variables.
The value of this methodology is contrasted considering four different publishing companies being representative as top sellers, mid-range sellers, the most varied, and one which sells a medium amount of different books.

The rest of this paper is structured as follows: next, Section \ref{sec:soa} presents a comprehensive review of the approaches found in the bibliography related to sales prediction in similar scopes. 
Then, Section \ref{sec:problem} introduces the problem of the estimation of print-run for new books, along with a description of the considered dataset.
Section \ref{sec:methodology} details the methodology considered in the study, from the data collection and preprocessing stages, to the selection of the forecasting methods to be applied in the experiments.  These are reported in Section \ref{sec:experiments}, which is also devoted to analyse the obtained results in feature selection and book sales forecasting using different datasets, with different amounts of features, for all the publishing companies. Obtained results for the four considered special cases are commented in Section \ref{sec:example_cases}. Finally, conclusions and future lines of work are presented in Section \ref{sec:conclusionsAndFutureWork}.

%********************************************************************************
\section{State of the art on sales forecasting} 
\label{sec:soa}

As far as we know, no previous attempt can be found in the literature 
%in which prediction methods have been applied to estimate the print run. 
in which prediction methods have been applied to estimate sales for new launched books.
However, similar techniques to those proposed in this paper, e.g. 
regression models \cite{Papalexopoulos1990}, 
neural networks \cite{Yoo1999} 
or fuzzy systems \cite{Mastorocostas2001}, 
have been applied to solve sales prediction problems in other industrial sectors.

Specifically, time-series prediction methods
\cite{Chu2003,Brown1959,Winters1960,Box1969,Papalexopoulos1990,KayacanUK10}  
is perhaps the most used technique to tackle sales forecasting
problems, although the efficiency of these techniques strongly depends on the 
field of application and the correctness of the problem data. However, since 
they require a large amount of data for predicting sales, these methods are not 
the most suitable for this task \cite{ChingChin2010}. 
Moreover, this kind of methods cannot be applied before the book is published, 
as is the objective of this paper. 
In this case, just a few variables are known in advance (pre-sales data)
\cite{ChingChin2010,FaderHardie2005,Madsen2008}, and some of 
them are categorical, which makes them harder to process, even more if 
the number of categories is high. As stated in those works, given these issues, 
finding adequate forecasting techniques and selecting the best method to use 
is also a problem.
An additional issue makes this a complex problem, since the predictions are 
influenced by external variables that must be taken into account, such as
seasonality, promotions, or fashions that expert managers might
subjectively apply \cite{Lapide1999,ChernWSF15}.
% [pedro] I have not found a reference. This is based on what Santiago told us.
% We can remove next sentence if you think it only should appear together with a reference.
For this reason, usually the methods used to predict book sales have
been generally based on experts' experience, who analyse data about sales and, 
taking into account their knowledge of industry, their experience, 
and their perception about trends, could make decisions on the companies 
production, i.e. how many books should be printed when a new book is published.

  %[pedro] sobre la industria editorial no hay trabajos sobre predicción de ventas.
  % Lo habitual ha sido usar el conocimiento de expertos para realizar predicciones 
  %  o aconsejar acerca de la cantidad de libros a imprimir.
  % Lo más parecido que he encontrado son dos trabajos (uno en realidad) de Moon et al. 
  %  sobre "Use Blog Information As Book Sales Prediction":
  %     Moon2010ICSSSM , Moon2010ICEC
Recently, some works have proposed analysing relationships between on-line 
information and book sales \cite{Moon2010ICSSSM,Moon2010ICEC}. 
Specifically, Moon et al. propose using the number of blog references as an 
indicator of the success of a book, and thus, its sales. 
Obtained results confirm relationships between the number of blog references 
and book sales, and at the same time different sales patterns are observed 
by analysing books in series (the first book published takes the longest period 
to achieve every single sales scale, while the second book takes the second 
longest period, and so on).

%
% pasar a contar los casos de la predicción de ventas en la 
% industria de la moda (SThomassey2014) y de nuevos productos (ChingChin2010)
%
% Son problemas muy similares, ya que no hay datos históricos de los nuevos productos,
% al igual que no los hay sobre ventas, y en los que además influyen mucho las variables 
% externas (estacionalidad, modas, fama del author, etc).
% Hay que echar mano de productos similares o simplemente de los datos que describen el nuevo producto.
%

In general, historical data are used to make time series, as well as for extracting common patterns and obtaining accurate predictions from them.
However, in the case of new products, such as the issue of new books, past patterns are difficult to observe as there is no available historical sales data 
\cite{ChingChin2010}. Thus, only pre-sales data are available in this problem, i.e. variables known before the book is on sale.

In order to deal with this issue, i.e. forecasting sales of new products, some authors have used different data mining methods, such as 
``Accurate Response'', an approach for demand estimation and production planning \cite{Hammond1990}, 
data clustering and fuzzy neural networks \cite{Chang2009} or 
extreme learning machine with adaptive metrics of inputs \cite{Xia2012}. 
%
In the same line, Thomassey et al. \cite{SThomassey2014} proposed using a 
clustering method together with decision trees for sales forecasting in 
textile-apparel fashion industry, a very similar issue to the new book 
publishing problem.
The same problem was previously faced by Xia et al. \cite{Xia2012} using a 
forecasting method based on extreme machine learning with adaptive metrics of inputs.

The complexity of this problem relies on the lack of historical data, on the 
short lifetime of the majority of items, and on the influence of variables such 
as promotions, fashions, or economic environment
 \cite{Thomassey2012,Xia2012,SThomassey2014}, as mentioned above.
%FERGU: completar con: "In our case, these ideas ..."
% [pedro] no sé a qué te refieres. Please, detalla un poco esto   :)

  %[Pedro] otro caso muy similar es el paper de ChingChin2010 sobre:
  % New tea product
  % New cosmetic product
  % New soft drink product
Ching et al. propose in \cite{ChingChin2010} a decision-support system for new 
product sales forecasting to solve three real-world sales forecasting 
problems, namely: new tea, cosmetics, and soft drink products. 
In that study, products are classified based on their sales patterns, so it is 
assumed that products in the same class would follow a similar sales pattern. 
However, as the authors state, the proposed system can not deal with qualitative 
data related to the products. Moreover, real-world cases, such as consumer 
electronics and fashion products, should be deeply analysed, as these industries 
present/launch new products every season.

  %[Pedro] otro caso a comentar es el paper de Chang2009 sobre: 
  % sales forecasting in printed circuit board industry.
In another hand, in those cases where historical data are available, classical time series forecasting methods have been applied, such as in \cite{Chang2009}, 
where a hybrid model integrating K-Means and fuzzy neural networks to forecast 
the future sales of a printed circuit board is proposed.
A similar method is described by Chern et al. \cite{ChernWSF15}, who analysed 
historical data throughout online reviews, reviewer characteristics, and review 
influences to understand how electronic `word-of-mouth' influences product sales. 
In this case, the proposed method is suitable for those sales forecasting problems 
with a big amount of online reviews and historical data.

Thus, taking into account the kind of problem we address in this
paper, i.e. the prediction of sales of a new product for which no
historical data are available, the classical forecasting methods based
on time-series are not adequate, although they can obviously be used once 
the book has been launched for sale. 
In our work the relevant variables in the prediction of sales are identified
by means of data visualisation and feature selection methods. 
Then, machine learning techniques are applied to predict sales and to infer 
the ideal print-run of a book.
In addition, since publishers need models to explain obtained predictions, 
`black box' forecasting methods may not be suitable, so models based
on decision rules have been tested as an alternative.

%********************************************************************************
\section{Problem description}
\label{sec:problem}

The problem to solve in this work is the forecasting of sales for new
launched books in the Spanish market, considering a set of `pre-sales
variables' or at least variables that are under the control of the
publisher before the book is going to be released to bookstores. The
objective is to offer a tool for publishers which will help them to
decide the best print run for a new book before its publication
process starts. This is a forecasting problem for which no specific
historical data are available and thus it is not simply a matter of
analyzing a time series, instead, we have considered data about other,
previously published, books and have taken as reference a set of
variables (or features), 
related with the book publication process.

The training dataset includes historical data about 6083 books
published in Spain for several years, with an initial set of 50
variables per pattern. It has been provided by the Spanish publishing
company Trevenque Editorial S.L. Table 
\ref{tabla:paramsOrig50} shows the initial set of variables describing
each book.

\begin{table*}[!ht]
\caption{Parameters describing a book, provided by the company Trevenque S.L.}
\label{tabla:paramsOrig50}
\begin{center}
\begin{tabular}{|c|c|}
\hline

\begin{minipage}{2.45in}\begin{small}
 \begin{verbatim}
reference
author
retail price
subject1
subject2
subject3
publisher
collection
bookbinding
print-run
total sales
dept stores sales
sales through delegates
rest of sales
total sales 1st year
mall sales 1st year
delegates sales 1st year
rest of sales 1st year
total distributed
distributed through dept stores
distributed through delegates
distributed - rest
total distributed 1st year
distrib. through dept stores 1st year
distrib. through delegates 1st year
\end{verbatim}
\end{small} \end{minipage}     &

\begin{minipage}{2.3in} \begin{small}
\begin{verbatim}
rest of distributed 1st year
reprints
number of reprints
total returns
returns - dept stores
delegates returns
rest of returns
total returns 1st year
returns -  dept stores 1st year
delegates returns 1st year
rest of returns 1st year
gifts
units distributed as novelty
total points of sale
points of sale - dept stores
delegates points of sale
rest of points of sale
total points of sale 1st year
dept stores points of sale 1st year
delegates points of sale 1st year
rest of points of sale 1st year
weeks for sale
positive sales environment
medium sales environment
negative sales environment
\end{verbatim}
\end{small} \end{minipage}    \\

\hline
\end{tabular}
\end{center}
\end{table*}

Starting from these raw data, a preprocessing method has been performed in order to remove incorrect or incoherent patterns, such as those with null values in required variables, or with negative sales, for instance. 
% (Paloma) Why remove those with null values and not to use an algorithm to fill those missing values?
% Antonio - because we should not fill important variables automatically, this adds noise to the data, so I think it is better to remove these patterns, since there are still a lot in the dataset. 
% Indeed, there were just a couple of patterns without important values, that probably were errors of the extraction process or bad insertions in the database.
Furthermore, the initial set of variables, presented in Table \ref{tabla:paramsOrig50}, 
has been reduced by an expert to use just those that can be known before a new book 
is launched, i.e. \textit{pre-sales data}. 
%Thus, their value do not depend on the book sales. 
The selected subset of features is described in Table \ref{tabla:params_pre_sales}.

\begin{table*}
% Shouldn't this table go to "Problem description"? It is a
% preliminary selection - JJ
% Antonio - done
\caption{Considered pre-sales features and their types. The third column shows
  the names of the variables used in the applied methods, and the last one indicates if they have been used as an input (independent) or an output (dependent) variable.} 
\label{tabla:params_pre_sales}
\begin{center}
\resizebox{12cm}{!}{
\begin{tabular}{|c|l|l|c|c|}
\hline 
No. & Feature Name & Variable & Type & In/Out\\
\hline 
1 & retail price (when launched) & \texttt{ret\_price} & numerical & input\\
2 & main subject (code) & \texttt{subject1} & numerical & input\\
3 &  publisher + collection (joint code) & \texttt{collection} & categorical & input\\
4 & bookbinding (code) & \texttt{binding} & categorical & input\\
5 & gifts (promotional books) & \texttt{gifts} & numerical & input\\
6 & units distributed as novelty & \texttt{distrib\_novelty} & numerical & input\\
7 & total number of points of sale & \texttt{tot\_points\_sale} & numerical & input\\
8 & total number of points of sale (1st year) & \texttt{tot\_points\_sale\_1st\_year} & numerical & input\\
9 & weeks on sale & \texttt{weeks\_sale} & numerical & input\\
10 & print run & \texttt{print\_run} & numerical & input\\
\hline 
\hline
11 & total sales & \texttt{total\_sales} & numerical & output\\
\hline 
\end{tabular}
}
\end{center}
\end{table*}

As it can be seen in this table, \texttt{total\_sales} is the output
variable, so it is the objective of the prediction methods. {\tt
  total\_sales} are a variable percentage of the {\tt print\_run}, and
thus there is obviously some correlation between them, which might
prevent us from using it to predict sales; it is quite obvious that
just increasing the print run, by itself, will not boost
sales. However, several of the other variables include implicitly a
certain size of the print run; for instance, you have to actually
print books to distribute them as novelty and at least one, for every
point of sale, so including the print run makes sense to get more
accurate estimates of the total sales that you are going to obtain out
of the total print run and check its influence on the overall sales
volume. Once this is included in the model and also as part of the
tree nodes, changing the print run might give the publisher an idea of
how many books are going to be actually sold out of that print run,
and taking that into account decide on the actual number of books that
are going to be printed. Just to be clear, printing books does not
actually sell them, but it is obviously necessary to print books
before you sell them, and once you have to do it, the publisher might
make decisions on printing a few more and check, with this model, what
would be the influence in sales if, for instance, you decide to send
more books to every point of sale or distribute more in novelty. We
will come back to this choice later on when examining the resulting
models.

Initially, there were three different features describing the subjects
and subsubjects of every book \textit{subject1,2,3} (see Table
\ref{tabla:paramsOrig50}), but for most of the patterns (books) there
is no value for subjects2 and 3, so we have discarded these features.

With regard to the feature \textit{publisher}, it has been removed from the final dataset since it is indeed included in the code of \texttt{collection}. In addition, we think it is not really relevant for the accuracy of the models and, moreover, publishers probably will not want to use the results of other publishing companies for predicting their own sales.
% Antonio - Think about this... should we mention this?
% (Paloma) If you finally leave it, please review it because this last sentence is a bit confusing
% Antonio - I guess it is now clearer

Therefore, in order to create simpler and more accurate prediction
models, and with the aim to provide them as a useful decision-aid tool
for the publisher, we have transformed the categorical variable
\texttt{subject1} into a numerical one, considering it refers to a
code of the Dewey Decimal Classification system for books
\cite{wiki:dewey}, and the existing similarity relationship between
the books inside a class or subclass. So, for instance, all the books
inside subdivisions of class 200, `Religion', are considered as
similar regarding their topics, so they could be also close in terms
of distance. The same reasoning can be applied to subclasses.

In this prediction problem there are other external variables that also affect 
the sales of a new book, such as those related to seasonality, trends, or 
promotions, that classical prediction methods can not deal with, since they are not included in the dataset due to the difficulty to model them. %FERGU: why?
% Antonio - explained.
However, an expert can consider them to modify the forecast obtained by 
the automatic method. Even publishing companies tend to define rules or 
indexes to subjectively adjust the results of prediction based on
external data. 
Thus, the proposed method will analyse the provided data and
variables and will yield a forecast of sales according to them. This
result could be used as a reference for a human expert, who would
consider it for his or her own prediction process.

% This section is too short and really does not describe the problem
% to its full extent. First, it should say if some of these variables
% were discarded for some reason. Group the variables we have in those
% available to publisher and those that can be used afterwards. And
% then give an idea of the complexity of the problem and how difficult
% it should be to find a solution - JJ
% Antonio - I have included much more contents form other sections. could you do the rest?

% You should include data visualization in problem description,
% because it shows the distribution of problem variables and then
% suggests ways to deal with it - JJ
% You should include in the section an exploratory data
% analysis that describes the range of variables, how many categories
% those variables have, and so an do forth - JJ
% Antonio - as you know SOM has been applied later as a feature selection method... I don't know if we could reference the figure of SOM here or the other way round (do the analysis here and reference later), but in that case, we should explain SOM before, which could be a mess...
% Antonio - If you meant using other visualization methods... could you do it also?

%********************************************************************************
\section{Methodology}
\label{sec:methodology}

There is not a standard methodology to solve the general problem of sales forecasting. However, the steps we have followed in this paper are those usually applied in a Knowledge Discovery in Databases (KDD) process \cite{fayyad1996data}. %These are described in the following sections.

Therefore, after the data has been collected and initially preprocessed (as described in previous section), other preprocessing techniques have been applied, such as Feature Selection (FS) \cite{kittler1986feature}, which is portrayed in Section \ref{subsec:data_preproces}. These methods check the relevance of all the variables of the patterns, aiming to reduce even more this set of features, in order to obtain intelligible models, in a lesser amount of time due to their lower complexity, and usually without accuracy loss.
This task may also help to identify leading indicators to be considered by a human expert to make predictions.
% You have to present the problem and not the solution. It's not "what
% you did last summer", it is SCIENCE. Why did you think FS was
% needed? What procedure did you use for it?
% Antonio - done
In this stage, both numerical and visualisation-based approaches have been applied, in order to select the best variables for the forecasting 
methods.  
% Antonio - Should we speak about 'forecasting' or 'regression', Sam? Coudl they be used with the same meaning?

Next, Section \ref{subsec:select_method} reports a selection of the best forecasting techniques that have been conducted, looking for those rule- or tree-based, together with numerical methods which could yield the best results.

Once the models and results have been obtained for the whole dataset (Section \ref{sec:experiments}), the same methodology has been applied to some specific cases of publishing companies in order to prove the value of the proposal (Section \ref{sec:example_cases}).
Some specific interpretable models have been created as a final result of the work. The aim is that these can be used as a decision-aid tool for the experts who have to decide on the best print-run quantity for new books to be published, in order to either maximise the benefit, or minimise the costs.

%-------------------------------------------------
%\subsection{Data collection}
%\label{subsec:data_collection}

%-------------------------------------------------
\subsection{Data preprocessing by feature selection}
\label{subsec:data_preproces}

At this stage, a Feature Selection (FS) process has been carried out, since it is an effective dimensionality reduction technique and an essential preprocessing method to discard less relevant or non-useful features \cite{Krishnapuram2004,Chen2011_FS_PTS}.

The aim of these algorithms is exploring the space of attributes to find out which subset of them yields the best results in a classification or prediction problem. To this end, the features describing every pattern are analysed. In this case, among the variables describing a new book, some may affect predictions as they provide information, while others may not be necessary and even redundant.

The application of FS also aims to reduce the complexity of the obtained 
forecasting models themselves, which, in turn, implies a reduction in the 
computational time for building these models. This has been a key 
factor in this work, because one of the objectives is to obtain `easily' 
interpretable decision models, which can be useful for human experts (publishers).
Regarding the algorithm, running time grows with the number of
features, making it impractical for problems with a large number of features 
\cite{Selvakuberan2008}; however, this is a secondary objective in this work.
%You need to lay out to the reader what you are going to do here. In
%order to reach objective x, we need to do y, that is why we use
%z. After using z, we prove that we have done y and thus reached
%objective x. For instance - JJ
% Antonio - Now it has been explained

% You have to justify its application to this particular problem. What
% was the runnning time with all variables? What reduction was your
% target? - JJ
% Antonio - I have explained before. The aim is not reducing the time, but the complexity of the obtained models in order to be more useful for a human

% During the evaluation, the set of all possible features is analyzed in order 
% to find the best set of features, ranking them according to some metric.
% In this process, the feature's predictive ability along with the degree 
% of redundancy between them is considered.

FS is one of the ways of performing a feature reduction, the other one is 
feature extraction \cite{kittler1986feature}. 
Feature extraction is a `transformation-based approach', therefore
it transforms the original meaning of the features. This method
involves creating a subset of new features by combining the
existing ones; thus it is employed when the semantics of the
original dataset will not be needed by any future process. On the
contrary, FS attempts to retain the meaning of the
original feature set. It is part of what is called semantic-preserving
techniques, known as `selection-based approaches' \cite{liu1998feature}. 
In the problem this work addresses, it is very important to maintain the 
semantics of initial variables set, and therefore just FS methods 
have been applied.

We have used some numerical FS methods included in Weka software \cite{Hall2009,Witten2011}. This tool separates the process into two parts:

\begin{itemize}
  \item Attribute evaluator: This is the method by which the feature subsets are assessed. In this work, both a correlation-based feature subset evaluation method ({\tt CfsSubsetEval}) and an adaptation of relief for attribute estimation ({\tt ReliefFAttributeEval}) evaluation method are used.

  The Correlation Feature Selection (CFS) method evaluates subsets of features/variables on the basis of the hypothesis made by Hall in \cite{Hall1998}, ``A  good  feature  subset  is  one  that  contains  features  highly  correlated  with
(predictive of) the class, yet uncorrelated with (not predictive of) each other''. This evaluator needs the numeric features to be transformed to nominal features first, by being discretized.%Pablo: the Oxford dictionary accepts both (zed and sed), but the one with Z is the one that appears as main in the entry
 The CFS evaluation function measures the ``merit'' of the subsets, which depend on the mean feature-class correlation, and the average feature-feature intercorrelation. In addition, the correlation is measured by three variations: the Minimum Description Length (MDL), the symmetrical uncertainty, and the ``relief''.

``Relief'' is the second evaluator used in this work. It is important to note that the relief method implemented in Weka is an updated version \cite{RobnikSikonja1997}, called RRELIEFF, of the original by Kira \& Rendell \cite{Kira1992}. The algorithm works by evaluating single attributes, and not whole subsets of attributes as CFS does. Also, it is applicable even working with both nominal and numeric features, without the need for transformations. For this algorithm, the weight of an attribute depends on the difference of an attribute value inside an instance with its neighbours, being these the nearest instance of the same class, and the nearest of the different class. Therefore, a good attribute is that which gives similar values for instances of the same class, and different values for instances of different class.
  \item Search method: This describes the structured way in which the set of possible feature subsets is studied, based on the results of the evaluator. With regard to the used search criteria, it is important to take into account that CFS evaluates attribute subsets, and Relief evaluates attributes separately.

For this reason, the search method used with CFS has been BestFirst, that searches the space of feature subsets by greedy hill-climbing augmented with a backtracking facility. On the other hand, Ranker is the searching criteria used with the Relief evaluator. This method ranks features by their individual evaluations.
\end{itemize}

In addition, in order to reduce the dimensionality of the input data
and to choose an appropriate set of variables, the Self-Organizing Map
(SOM) \cite{kohonen1998} method has also been applied as a `subjective' (based on expert's opinion) FS technique based on the visualisation of the data as a set of 2D planes (one per feature of the dataset), and their interpretation for finding out relationships between these features.

% Why? Is that one of the two forms of feature selection? What do we expect to learn from it? - JJ 
% Antonio - explained here and below. Please check.

%--- new SOM description % -Sam

SOM is a type of unsupervised neural network inspired by the
topographical organisation of the sensory cortex of human brain
\cite{kohonen1998}. It maps highly dimensional data into a lower
dimensional space where similar data patterns are located in more
adjacent areas on the grid. 
% It is a very useful analysis tool for visualising complex data with high number of variables in a simple geometric plane. 
%Therefore, it could be used for different tasks such as feature selection and data clustering and visualisation. %Pablo: some cite here?
This method is normally used as a data visualization tool, since it allows to project any multi-variated dataset in a two-dimensional display, so an expert can detect topological relations among its neurons and thus make inferences or find interesting relationships on the input dataset (such as clusters, for instance).

Basically, SOM consists of two layers: input and output layer, also
known as Kohonen, layer. Input layer is fully connected with the 
Kohen layer which is usually a two dimensional grid of neurons
(i.e. units). This means that every neuron is linked to every input in
the input layer.

SOM follows an iterative process applied in two stages: First, the
weights of the network are initialised randomly, then the n-dimensional
input vectors are presented to the network sequentially so the
distance between the weights of the neurons and the inputs are
computed using the simple Euclidean distance; in the second stage, the
neuron which has the shortest distance with the presented input vector
is updated along with the neurons in its neighbourhood. This neuron is called the best matching unit (BMU). The goal of the learning process is to
minimise the neighbourhood distance in the output clusters.

Once the SOM learning process is finished, a visual representation technique, named the Unified distance matrix (U-matrix) \cite{UmatUlts}, has been applied. This method uses the obtained codevectors (vectors of variables of the problem) as data source, and generates a matrix where each component is a distance measure between two adjacent neurons. In the U-matrix, the distance between adjacent neurons is measured and assigned different colours. Different colours between neurons indicate large distances (boundaries) while similar colours represent close areas (clusters).

% Antonio - a better description of U-Matrix. ;D

%---

This method is usually applied to analyse data as similar items tend to be 
mapped close together, while those items dissimilar, are mapped apart.
Also, the SOM graph represents and highlight very clearly those
regions, or clusters, with high training sample concentration and fewer
where the samples are scarce. U-Matrix graph, together with the projection of variables into separate planes, the so-called component planes analysis (that SOM yields), will aid an expert to identify the most relevant features on every cluster, i.e. those with a higher influence on the creation of that cluster.
%FERGU: mencionarlo en la intro
% Which is interesting because... JJ
% Antonio - explained. ;)

Finally, Section \ref{sec:experiments} provides a comparison between the feature reduction through SOM, and the feature selection techniques that have been described.

%-------------------------------------------------
\subsection{Selecting an appropriate forecasting method}
\label{subsec:select_method}

As previously stated, there are many prediction methods available in the 
literature, each one with a different parameter set that may affect the 
obtained results.

However, as in the case of new books only some descriptive data are available,
compared to those cases in which historical data on sales of published books 
is available, not all forecasting methods can be used in this specific problem.

Thus, in this research five forecasting methods, based on the State of the Art literature 
review (see Section \ref{sec:soa}) 
\cite{Madsen2008,ChingChin2010,Thomassey2012,Xia2012,SThomassey2014}, have been used. 
Well known forecasting methods implemented in the Weka tool have been chosen, as these methods are widely known, and could be very helpful for practitioners to reproduce experiments and even to solve similar sales prediction problems using
these methods. % Why these and not others? Exploratory data analysis?
               % - JJ

Specifically, the following forecasting methods are proposed: 
        % Proposed why? - JJ
        % because those methods were those that could lead with the problem data - pedro
\begin{itemize}

 % weka.classifiers.trees.M5P
 %		http://weka.sourceforge.net/doc.dev/weka/classifiers/trees/M5P.html
 \item \emph{M5 Model trees (M5P)}: 
A decision tree consists of answer-nodes, that indicate a class, and decision-nodes, 
that contain an attribute name and branches to other sub-decision trees.
Building a decision tree can be done using many algorithms, i.e. ID3 and C4.5 \cite{Quinlan1986}.
However, in order to use this model in regression problems, some authors have 
extended the model using methods such as the M5 model tree \cite{Quinlan1986,Quinlan1992,Wang1997,WittenFrank2000}, 
by combining a conventional decision tree and generating linear regression 
functions at the nodes.

The construction of a model tree is similar to that of classical decision 
trees \cite{Solomatine2004}:
The process breaks the input space of the training data through decision points 
(nodes) to assign a linear model suitable for that sub-area of the input space. 
This process may lead to a complex model tree.

Model tree models can learn and tackle tasks with high dimensionality, even up to 
hundreds of attributes, and generate reproducible and comprehensible representations, 
what makes them potentially more successful in the eyes of decision makers.

The final model consists of the collection of linear sub-models that brings the 
required non-linearity in dealing with the regression problem
and both the predicted values at the answer-nodes along with the path from the 
root to that node is given as a result.

 % weka.classifiers.lazy.IBk 
 %		http://weka.sourceforge.net/doc.dev/weka/classifiers/lazy/IBk.html
 \item \emph{k-Nearest Neighbours (kNN)}: 
The kNN \cite{Aha1991,Mitchell1997} method is an instance-based classifier in 
which the unknown instances are classified by relating them to the known 
instances using a distance measurement/function: Euclidean, Minkowsky or minimax.

The main idea behind the algorithm is that two instances far enough in the space, 
taking into account the distance function, are less likely to belong to the 
same class than two closely situated instances.

The classification algorithm locates the nearest neighbour in instance space and 
assigns the class of that neighbour to the unknown instance.
In order to improve the robustness of the model, several neighbours can be 
located, assigning the resulting class to an unclassified vector using the 
closest $k$ vectors found in the training set by majority vote.

 \item \emph{Random Forest (RF)}: This method is based on the
   construction of a set of decision trees using a stochastic process
   over the basis of C4.5 algorithm. It was proposed by Breiman
   \cite{Breiman2001} and aims to create independent and uncorrelated
   trees based on different and random input vectors, following the
   same distribution. 
The result is the average of the generated trees during the process.

 % weka.classifiers.functions.LinearRegression
 %		http://weka.sourceforge.net/doc.dev/weka/classifiers/functions/LinearRegression.html
 \item \emph{Linear Regression (LR)}:
 Linear Regression \cite{Cohen2003,Yan2009,Rencher2102,EnkeT05} method models the 
relationship between a scalar dependent variable (outcome) and several independent 
variables (predictors) over the range of values in the dataset. 
In this statistical technique, data are modelled using linear functions to estimate unknown model parameters, examining the linear correlations between independent and dependent variables.

This method of regression provides an adequate and interpretable description of 
how the input affects the output as it models the dependent variable as a linear 
function of the independent variables. Moreover, the linear relation can be 
solved using the least squares method, that minimises the error between the 
current data and the regression line \cite{McClendon2015}.

 % weka.classifiers.functions.SMOreg
 %		http://weka.sourceforge.net/doc.dev/weka/classifiers/functions/SMOreg.html
 \item \emph{Support Vector Machine for Regression (SVM)}:
Support vector machine (SVM) \cite{Cortes1995,Shevade1999,MinSVM05} method is based on  
statistical learning theory and has successfully used in classification and  
regression problems \cite{Cao2003,Jari2008}.

In classification problems the algorithm searches for an optimal hyperplane 
that separates two classes, maximising the margin between them. 
In the case of regression the algorithm chooses a hyperplane close to as many 
of the data points as possible, minimising the sum of the distances from 
the data points to the hyperplane. 
In both cases, the hyperplane is defined by a subset of training set samples, 
called support vectors.
%This method works well even if the space is highly dimensional and the problem is not linearly separable.

\end{itemize}

The objective is finding the most appropriate forecasting method that minimises 
the error between the obtained forecast and the current sales for each new book 
published.

\subsection{Models evaluation}

At this step, the developed forecasting models are evaluated and assessed based on different evaluation criteria. In addition to the \textit{correlation coefficient}, we have considered the widely accepted standard formulae in the literature \cite{gepsoft} (Equations \ref{eq:MAE} to \ref{eq:RRSE}), which are implemented in tools like Weka \cite{pentaho,Witten2011} or R \cite{otexts,Hyndman2013}.
Specifically, in order to compute the forecasting errors of the different methods, Mean absolute error (MAE), Root mean squared error (RMSE), Relative absolute error (RAE) and Root relative squared error (RRSE) have been selected:

\begin{itemize}
  \item \emph{Mean Absolute Error} (MAE):
        \begin{equation}\label{eq:MAE}
            MAE = \frac{1}{n}\sum_{i=1}^n {\mid p_i - o_i\mid}
        \end{equation}

  \item \emph{Root Mean Squared Error} (RMSE):
        \begin{equation}\label{eq:RMSE}
            RMSE = \sqrt{ \frac{1}{n}\sum_{i=1}^n {(p_i - o_i)}^2 }
        \end{equation}

  \item \emph{Relative absolute error} (RAE):
        \begin{equation}\label{eq:RAE}
            RAE = \frac{ \sum_{i=1}^n {\mid p_i - o_i\mid} }{ \sum_{i=1}^n {\mid \bar{o} - o_i\mid} }
        \end{equation}

  \item \emph{Root relative squared error} (RRSE):
        \begin{equation}\label{eq:RRSE}
            RRSE = \sqrt{ \frac{ \sum_{i=1}^n {(p_i - o_i)}^2  }{ \sum_{i=1}^n {(\bar{o} - o_i)}^2 }  }
        \end{equation}
\end{itemize}
where $o_i$ is the current value for the instance $i = {1,...,n}$, $\bar{o}$ is the mean of $o$ and $p_i$ is the obtained prediction.

%********************************************************************************
\section{Experiments and Results}
\label{sec:experiments}

% An intro should go here - JJ
% Antonio - done

This section presents the experiments conducted over the dataset considering the pre-sales variables forecasting new book sales.
Feature selection techniques have also been applied in order to test their utility by comparing the obtained results.

We have analysed the whole dataset (all the publishing companies), but also performed some experiments considering specific and representative publishing companies as example cases.
Some of the obtained decision trees have been plotted to show their simplicity and to remark their real value as a decision-aid tool for publishers.

%--------------------------------------------------------------------------

\subsection{Experimental setup}

In most data mining and machine learning algorithms, the values of
some parameters have a high influence on its performance. We have conducted a systematic experimentation process, along with a revision of the bibliography describing the best configuration parameters for the methods we have applied here.

Thus, in our experiments, the best value of $k$ in the $k$-nearest neighbour was empirically found at $k=3$. % What???? - JJ
For SVM, the widely used Radial Basis Function (RBF) kernel is selected. As reported in several studies, RBF kernel has a reliable performance and can analyse higher dimensional data \cite{yu2004ec,chao2015construction}. Two important parameters should be tuned in SVM: the Cost $(C)$ and Gamma $(\gamma)$ of the RBF kernel. We apply grid search which is the most common approach
to find $C$ and $\gamma$. The best performance of SVM was found at
$C=200$ and $\gamma = 0.01$. For MLP, we apply it with one
hidden layer since it was proven in the literature that an MLP with
one hidden layer can approximate any continuous or discontinuous
function \cite{hornik1989multilayer,mirjalili2014let}. However, there
is no rule of thumb for selecting the best number of neurons in the
hidden layer. One suggested method that is followed in this work is to
set the number of neurons in the hidden layer to $2 \times I + 1$
where $I$ is the number of inputs in the dataset
\cite{wdaa2008differential,mirjalili2014let,mirjalili2015effective}.

All experiments have been performed using the standard 10-fold cross validation \cite{Geisser1993,Kohavi1995,Devijver1982}: then, the dataset is partitioned into 10 parts where the training process is carried out on 9 parts and after this, the resulting model is tested on the left 10th part. This process is repeated 10 times, training the models using different 9 parts. At the end of the process the 10 evaluation results on the testing parts are averaged. 
This technique is used to estimate how accurately the predictive model will perform in practise, limiting, at the same time, the overfitting problem.

In addition, due to the stochastic component present in some of the methods and also in order to fairly evaluate the running time, 30 repetitions of each experiment have been done per method. Then the average and standard deviation have been computed.

All experiments have been conducted on an  Intel(R) Core(TM) i5-2400 CPU at 3.10Ghz with a memory of 4 GB running Windows 7 professional 32bit operating system and Java JRE\_1.7.0\_72.

%--------------------------------------------------------------------------

\subsection{Results for all publishers}
\label{subsec:results_all_publishers}

In the first stage of the experiments, forecasting methods have been
applied to the whole dataset. All the pre-sales features have been
considered, including the categorical ones, which makes it harder to
process the input patterns and build the different models. SVM for
instance, must previously transform these categories into binary
variables in order to deal with the data, which means the creation of
a different (and new) variable per each possible value. Thus, in the
case of \emph{publisher} or \emph{collection} this would imply the definition of
hundreds of new variables.

Correlation coefficient, error measures and time to build the model have been computed for every method. Obtained results are presented in Table \ref{tab:all_publishers_no_fs}.

\begin{table*}
\caption{Predicting Total sales for all publishers (all the features). 30 repetitions, 10-fold cross validation. Best values in bold.
\label{tab:all_publishers_no_fs}}
\centering{}%
\scalebox{0.78}{

\begin{tabular}{|c|l|l|l|l|l|l|}
\hline 
 & $R^{2}$  & MAE  & RMSE  & RAE (\%)  & RRSE (\%)  & Time (s)\tabularnewline
\hline 
M5P  &  \textbf{89.0 $\pm$ 0.09}  &  \textbf{214.37 $\pm$ 22.96}  &  \textbf{634.23 $\pm$ 307.76}  &  \textbf{37.46 $\pm$ 3.12}  &  \textbf{48.32 $\pm$ 25.16} & 18.85 $\pm$ 0.61\tabularnewline
\hline 
kNN  & 78.7 $\pm$ 0.08  & 272.29 $\pm$ 32.38  & 840.59 $\pm$ 240.11  & 47.48 $\pm$ 3.39  & 63.07 $\pm$ 10.29  &  \textbf{0.001 $\pm$ 0.01}\tabularnewline
\hline 
RF & 84.0 $\pm$ 0.05  & 253.37 $\pm$ 29.43  & 788.94 $\pm$ 282.01  & 44.16 $\pm$ 2.76  & 57.77 $\pm$ 8.15  & 4.91 $\pm$ 0.12\tabularnewline
\hline 
LR & 87.2 $\pm$ 0.07  & 308.68 $\pm$ 25.44  & 680.91 $\pm$ 226.41  & 54.05 $\pm$ 4.38  & 52.13 $\pm$ 18.29  & 984.68 $\pm$ 128.21\tabularnewline
\hline 
SVM  & 87.6 $\pm$ 0.04  & 248.76 $\pm$ 27.49  & 732.64 $\pm$ 260.56  & 43.37 $\pm$ 2.35  & 53.47 $\pm$ 6.29  & 6169.24 $\pm$ 281.49\tabularnewline
\hline 
\end{tabular}
}
\end{table*}

As far as the different methods are concerned, M5P model obtained the best results, both in time and error. The advantage of M5P is that it can learn and tackle classification tasks that have high dimensionality and categorical variables efficiently. Moreover, M5P generates reproducible and comprehensible representations that provide information on the decision process and does not require human intervention either for the operation or for interpretation. Thus, it is very suitable model be incorporated in a publishing company 
processes (as we will show in Section \ref{subsec:decision_trees}).

Taking into account the obtained results by SVM model, this method obtains comparable errors to those of M5P. Even when using roughly tuned parameter values, it shows a good generalisation performance. However, SVM is much slower than the others to build the model, not only when using the dataset with categorical input variables, such as \emph{collection} and \emph{binding}, but also when using the datasets with numerical variables, not to mention the extra time needed to optimize its hyperparameters.

According to Table \ref{tab:all_publishers_no_fs}, it can also be noticed that the Linear Regression model outperforms some other complex methods like the Random Forest, both in time and in obtained errors. Although, Linear regression is considered as a simple model to build, more powerful models can obtain better results in comparable time.

kNN is the fastest method used in this study in terms of training time, as it 
is the simplest one. However, the model is slow when classifying large number of 
training examples, and it exhibits a poor generalisation ability since it does not learn anything from the training data.

%Moreover, in these experiments it has been seen the difficulty for some classifiers like the SVM to process information related to categorical variables, such as \emph{collection} and \emph{binding}.

As a first conclusion, paying attention both to the cost in time to obtain the forecasting models and their accuracy, the M5P model would be the most adequate to solve the problem of how many books should be printed when a new book is published. This forecasting method, with these additional advantages, can be used for products with historical sales data and for new products with little or no historical data available.

%--------------------------------------------------------------------------

\subsection{Feature selection}
\label{subsec:feature_selection}

In the second stage of the experiments, different feature selection methods have been applied. The aim of these methods is to reduce the number of features considered in developing the forecasting models which consequently simplify the complexity of the models, improve their interpretability and decrease the training time. The simpler models generated after performing feature selection could give the decision maker a better insight on the underlying process by identifying the most influencing variables.

We have used both the {\tt CfsSubsetEval} \cite{Hall1998} (a correlation-based feature subset evaluation method) and the ReliefFAttributeEval \cite{RobnikSikonja1997} (an adaptation of relief for attribute estimation) evaluation methods, implemented in the Weka tool and described in Section \ref{subsec:data_preproces}.

The dataset has also been analysed using SOM, and a visual exploration of the obtained results has been performed in order to determine the best variables to consider.
An important preprocessing step has been carried out before conducting the SOM analysis. Thus, the two categorical features \emph{collection} and \emph{binding} have been removed from the analysis since SOM cannot deal with them without a previous transformation into numerical ones. However, since they have a large number of unique different values this would be finally impracticable and useless.

In the analysis, several configurations for SOM have been tested, and the best one regarding both the \textit{quantisation error} and \textit{topographic error} has been chosen. The first error is a measure which represents the average distance between every input pattern and its BMU (i.e. how good the map fits the dataset). The latter refers to the topology of the map and its preservation, so it measures an average value computed looking at both, the first and the second BMUs for every pattern. If they are neighbours (directly connected in the map) a `1' is considered, otherwise a `0' is taken. A small value would mean that the map preserves well its topology.

Thus, the chosen map to represent the data has $26 \times 16$ neurons and an hexagonal neighbourhood topology.  The best SOM training results have been obtained with a final quantisation error of 0.691 and topographic error of 0.046.

Figure \ref{fig:componentplanes} shows the U-matrix along with the component planes analysis resulted from SOM training. The term component refers in SOM to a variable or feature of the problem.
The component planes in Figure \ref{fig:componentplanes} show the trend of values of the prototype vectors of the SOM map units. These values are represented as colours. The colourbar on the right of each component plane shows the indication of each colour. On the other side, the U-matrix on the top
left corner shows the distances between the adjacent units. It is
important to note that high values in the matrix indicates large
distance between the units of the map. It can be noticed also that
there are more hexagons in the U-matrix than component planes since
the distances between units are also represented as inter-neuron hexagons.

Examining the component planes shown in Figure \ref{fig:componentplanes}, it can be noticed that some variables have very similar colour maps which means that they are highly correlated. In this case, we have two examples; the first is the \emph{distrib\_novelty} and \emph{print\_run}variables while the second is the \emph{tot\_points\_sale} and \emph{tot\_points\_sale\_1st\_year}. In order to make the dataset having less number of uncorrelated features with each other, one of the redundant features is removed from each pair of features. Therefore, the \emph{distrib\_novelty} and \emph{tot\_points\_sale\_1st\_year} were omitted from the dataset.

\begin{figure*}[ht]
\begin{center}
\includegraphics[scale=0.28]{som_planes.eps}
\end{center}
\caption{U-matrix and component planes.}
\label{fig:componentplanes}
\end{figure*}

The selected features using each one of the three methods are shown in Table \ref{tab:features_selected}. Logically, just the independent or input variables are considered.

\begin{table*}
\caption{Input features/variables selected by each method
\label{tab:features_selected}}
\centering{}%
\begin{tabular}{|c|c|c|c|c|}
\hline 
No. & Variable & ReliefFAttributeEval & CfsSubsetEval & SOM\\
\hline 
1 & \texttt{ret\_price} & $\checkmark$ &  & $\checkmark$\\
\hline 
2 & \texttt{subject1} & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\hline 
3 &  \texttt{collection} &  &  & \\
\hline 
4 & \texttt{binding} &  &  & \\
\hline 
5 & \texttt{gifts} & $\checkmark$ &  & $\checkmark$\\
\hline 
6 & \texttt{distrib\_novelty} & $\checkmark$ &  & \\
\hline 
7 & \texttt{tot\_points\_sale} & $\checkmark$ &  & $\checkmark$\\
\hline 
8 & \texttt{tot\_points\_sale\_1st\_year} & $\checkmark$ &  & \\
\hline 
9 & \texttt{weeks\_sale} & $\checkmark$ &  & $\checkmark$\\
\hline 
10 & \texttt{print\_run} & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\hline 
\end{tabular}
\end{table*}

After applying the feature selection process, we have run the same
forecasting models as in Section \ref{subsec:results_all_publishers}
for the whole dataset (all the publishers), but considering only the
selected features by each method. The results are shown in Table
\ref{tab:results_all_publishers_fs}. Examining the obtained results
for each feature selection method, it can be seen that RFE got the
best results compared to SOM and CSE. It can be noticed also that we
obtain better performance in M5P, kNN and RF using only features
selected by RFE than using all features. %Better publish results in
                                %the same table or refer to the table
                                %where it was published. Is this
                                %difference statistically significant?
                                %- JJ
As expected, in general, all feature selection methods have reduced the training time needed to build the forecasting methods. This is because they reduced the considered features and mainly because they excluded the categorical features, which are the hardest to be managed (they have many possible values). On the other side, CSE and SOM did not perform as good as RA. This could be explained due to the extensive reduction made by CSE and due to the `subjective' (but not so accurate) human intervention in the case of SOM.

\begin{table*}
\centering{}%
\caption{Predicting Total sales for all publishers based on features selected
by each method. Best values in bold.
\label{tab:results_all_publishers_fs}}
% RELIEFATTRIBUTEEVAL
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
\multicolumn{7}{|c|}{\textit{ReliefAttributeEval}}\\
\hline 
 & $R^{2}$  & MAE & RMSE & RAE (\%) & RRSE (\%) & Time (s)\\
\hline 
M5P & 89.2 $\pm$0.10 & 211.15 $\pm$22.41 & 616.35 $\pm$321.25 & 36.93 $\pm$3.41 & 47.30 $\pm$26.83 & 0.49 $\pm$0.02\\
\hline 
kNN & 86.9 $\pm$0.09 & 231.78 $\pm$24.25 & 668.03 $\pm$205.14 & 40.48 $\pm$3.06 & 51.28 $\pm$16.81 &  \textbf{0.002 $\pm$0.004}\\
\hline 
RF &  \textbf{91.0 $\pm$0.05} &  \textbf{198.16 $\pm$19.50} &  \textbf{589.35 $\pm$195.80} &  \textbf{34.61 $\pm$2.27} &  \textbf{44.21 $\pm$11.63} & 3.63 $\pm$0.20\\
\hline 
LR & 86.4 $\pm$0.08 & 346.88 $\pm$25.40 & 695.90 $\pm$242.49 & 60.75 $\pm$4.55 & 53.16 $\pm$19.43 & 0.01 $\pm$0.01\\
\hline 
SVM & 84.2 $\pm$0.04 & 252.87 $\pm$33.38 & 858.26 $\pm$318.23 & 44.01 $\pm$3.04 & 62.22 $\pm$7.15 & 38.97 $\pm$5.75\\
\hline 
\end{tabular}
}
% CFSSUBEVAL
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
\multicolumn{7}{|c|}{\textit{CfsSubEval}}\\
\hline
 & $R^{2}$  & MAE & RMSE & RAE (\%) & RRSE (\%) & Time (s)\\
\hline 
M5P &  \textbf{87.0 $\pm$0.09} &  \textbf{266.78 $\pm$23.12} &  \textbf{662.32 $\pm$245.27} &  \textbf{46.66 $\pm$3.25} &  \textbf{50.9 $\pm$20.35} & 0.41 $\pm$0.04\\
\hline 
kNN & 84.0 $\pm$0.10 & 289.09 $\pm$22.83 & 743.92 $\pm$260.45 & 50.59 $\pm$3.50 & 57.30 $\pm$22.90 &  \textbf{0.001 $\pm$0.001}\\
\hline 
RF & 85.0 $\pm$0.11 & 278.42 $\pm$23.07 & 714.13 $\pm$278.65 & 48.71 $\pm$3.43 & 55.26 $\pm$24.41 & 2.48 $\pm$0.13\\
\hline 
LR & 85.9 $\pm$0.06 & 356.35 $\pm$22.43 & 708.06 $\pm$195.15 & 62.42 $\pm$4.22 & 53.97 $\pm$14.05 & 0.01 $\pm$0.001\\
\hline 
SVM & 86.0 $\pm$0.06 & 288.47 $\pm$30.32 & 817.56 $\pm$253.24 & 50.30 $\pm$2.38 & 60.27 $\pm$5.68 & 32.73 $\pm$5.74\\
\hline
\end{tabular}
}
% SOM
\centering{}% SOM
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
\multicolumn{7}{|c|}{\textit{SOM}}\tabularnewline
\hline 
 &  $R^{2}$  & MAE  & RMSE  & RAE (\%)  & RRSE (\%)  & Time (s)\tabularnewline
\hline 
M5P  & 86.6 $\pm$0.10  & 234.09 $\pm$23.19  & 699.56 $\pm$317.20  & 40.92 $\pm$3.25  & 53.70 $\pm$26.87  & 0.475 $\pm$0.12\tabularnewline
\hline 
kNN  & 86.8 $\pm$0.09  & 237.91 $\pm$25.34  & 665.64 $\pm$212.05  & 41.55 $\pm$3.14  & 50.96 $\pm$16.72  &  \textbf{0.001 $\pm$0.004}\tabularnewline
\hline 
RF &  \textbf{89.8 $\pm$0.05}  &  \textbf{211.87 $\pm$20.07}  &  \textbf{617.78 $\pm$187.16}  &  \textbf{37.02 $\pm$2.46}  &  \textbf{46.58 $\pm$11.14}  & 2.93 $\pm$0.11\tabularnewline
\hline 
LR  & 85.5 $\pm$0.08  & 351.99 $\pm$23.72  & 714.76 $\pm$214.21  & 61.65 $\pm$4.28  & 54.67 $\pm$16.88  & 0.01 $\pm$0.01\tabularnewline
\hline 
SVM  & 84.5 $\pm$0.04  & 267.71 $\pm$33.82  & 888.33 $\pm$308.93  & 46.60 $\pm$2.87  & 64.62 $\pm$5.45  & 32.27 $\pm$4.74\tabularnewline
\hline 
\end{tabular}
}
\end{table*}

%********************************************************************************

\section{Experiments of special publishing companies cases}
\label{sec:example_cases}

This section presents further experiments conducted on data concerning four interesting special publishing company cases. These publishing companies were selected as representative cases which could be very interesting for decision makers and managers in the domain. The four cases are described as follows:

\begin{itemize}
    \item Best seller publishing company: this case represents the publishing company that sells the highest number of copies regardless the title of the books sold. Therefore, it is an example of a case on a higher business scale.
    \item Mid-range seller publishing company: this represents a publishing company that sells medium number of copies compared to other publishing companies (Medium business scale). 
    \item Highest variation publishing company: this is an example of a publishing company that follows a strategy in publishing a very wide range of different titles.
    \item Medium varied publishing company: this case represents a publishing company that has a median rank between the other publishing companies regarding the variety of titles sold.
\end{itemize}

Table \ref{Table:special_cases} shows the number of sold copies and the number of titles published by the selected companies.

\begin{table*}
\caption{Description of publishing companies considered as special cases.}
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|l|l|l|}
\hline 
 Case & Different books sold & Total copies sold & Description\\
\hline 
\textit{All publishing companies} & 6083 & 3201645 & 209 publishing companies\\
\hline 
\textit{UDL929} & 216 & 460752 & Best selling publishing company\\
\hline 
\textit{UDLW12} & 97 & 2635 & Mid-range selling publishing company\\
\hline 
\textit{UDL704} & 293 & 58236 & publishing company with most different\\ 
       &     &       & books sold\\
\hline 
\textit{UDLU11} & 180 & 249481 & publishing company with medium number\\ 
       &     &       &  of different books sold\\
\hline 
\end{tabular}
}
\label{Table:special_cases}
\end{table*}

The experiments applied on the selected special cases follow the same methodology used previously for the whole dataset. First, the forecasting models are applied on each dataset using all features, then, the three feature selection methods (RFE, CSE and SOM) are used to reduce the set of features in each case. Second, the classification models are applied again on the reduced datasets to evaluate their performance.

Evaluation results of the developed models (without performing any feature selection method) for the selected four cases are shown in Tables \ref{Table:case1results}, \ref{Table:case2results}, \ref{Table:case3results} and \ref{Table:case4results}. The results in these tables show that the classifiers performed similarly to the case of the global models for all publishers which were previously presented in Table \ref{tab:all_publishers_no_fs}. In general, M5P, RF and SVM performed better than the `simple' classifiers kNN and LR. In the case of Best selling publisher houses (i.e. UDL929) and publishers who publish wide range of different books (i.e. UDL704), M5P could be recommended as it shows more accurate forecasting results than the other models. While in the case of medium sized publishers, in terms of sales and number of published books, the RF model would be adequate.

\begin{table*}
\caption{Predicting Total sales for publishing company-UDL929 (The best selling publishing company). Best values in bold. }
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE  & RMSE  & RAE (\%)  & RRSE (\%)  & Time (s)\tabularnewline
\hline 
\hline 
M5P  &  \textbf{90.0 $\pm$0.10} &  \textbf{670.24 $\pm$181.05}  &  \textbf{1065.34 $\pm$413.84}  &  \textbf{40.83 $\pm$10.86}  &  \textbf{43.56 $\pm$19.87}  & 0.03 $\pm$0.002\tabularnewline
\hline 
kNN  & 73.3 $\pm$0.19  & 1057.21 $\pm$385.04  & 1875.85 $\pm$916.96  & 61.88 $\pm$12.37  & 68.75 $\pm$15.98  &  \textbf{0 $\pm$0.0001}\tabularnewline
\hline 
RF  & 87.6 $\pm$0.11  & 771.83 $\pm$326.34  & 1490.49 $\pm$930.58  & 45.40 $\pm$14.14  & 55.91 $\pm$31.70  & 0.08 $\pm$0.004\tabularnewline
\hline 
LR  & 86.3 $\pm$0.12  & 922.33 $\pm$289.01  & 1399.84 $\pm$619.62  & 56.11 $\pm$16.17  & 56.43 $\pm$25.09  & 0.005 $\pm$0.001\tabularnewline
\hline 
SVM  & 86.7 $\pm$0.1  & 873.22 $\pm$345.16 & 1621.48 $\pm$985.24  & 50.85 $\pm$11.83  & 57.23 $\pm$20.40  & 0.04 $\pm$0.016\tabularnewline
\hline 
\end{tabular}
}
\label{Table:case1results}
\end{table*}

\begin{table*}
\caption{Predicting Total sales for PublishingCo-UDLW12 (The median selling publishing company). Best values in bold. }
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE  & RMSE  & RAE (\%)  & RRSE (\%)  & Time (s)\tabularnewline
\hline 
\hline 
M5P  & 83.2 $\pm$0.24  & 14.18 $\pm$6.62  & 23.82 $\pm$11.86  & 43.41 $\pm$20.39  & 55.79 $\pm$30.47  & 0.009 $\pm$0.002\tabularnewline
\hline 
kNN  & 79.3 $\pm$0.24 & 15.34 $\pm$8.71  & 26.57 $\pm$16.64  & 46.20 $\pm$23.77  & 60.09 $\pm$35.21  &  \textbf{0.0001 $\pm$0.0002}\tabularnewline
\hline 
RF &  \textbf{85.8 $\pm$0.17}  & 14.19 $\pm$6.69  &  \textbf{23.57 $\pm$11.83}  & 42.53 $\pm$16.80  & 53.30 $\pm$23.41  & 0.03 $\pm$0.004\tabularnewline
\hline 
LR  & 82.1 $\pm$0.25 & 14.62 $\pm$7.92  & 24.79 $\pm$15.08  & 44.57 $\pm$24.27  & 56.85 $\pm$35.07  & 0.002 $\pm$0.0005\tabularnewline
\hline 
SVM  & 83.5 $\pm$0.21  &  \textbf{14.01 $\pm$7.61}  & 25.61 $\pm$15.30  &  \textbf{40.66 $\pm$16.03}  &  \textbf{53.28 $\pm$18.45}  & 0.02 $\pm$0.02\tabularnewline
\hline 
\end{tabular}
}
\label{Table:case2results}
\end{table*}

\begin{table*}
\caption{Predicting Total sales for PublishingCo704 (highest number of different titles sold). Best values in bold.}
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 &  $R^{2}$  & MAE  & RMSE  & RAE (\%)  & RRSE (\%)  & Time (s)\tabularnewline
\hline 
\hline 
M5P  &  \textbf{88.9 $\pm$0.07} &  \textbf{71.26 $\pm$21.29}  &  \textbf{125.19 $\pm$50.09}  &  \textbf{39.25 $\pm$11.77}  &  \textbf{47.50 $\pm$15.93}  & 0.05 $\pm$0.03\tabularnewline
\hline 
kNN  & 82.6 $\pm$0.10  & 87.60 $\pm$23.21  & 158.63 $\pm$54.16  & 47.62 $\pm$10.32  & 59.17 $\pm$14.43  &  \textbf{0.0001 $\pm$0.0003}\tabularnewline
\hline 
RF & 87.1 $\pm$0.08  & 74.63 $\pm$22.17  & 139.74 $\pm$54.36  & 40.62 $\pm$10.21  & 52.22 $\pm$16.27  & 0.11 $\pm$0.01\tabularnewline
\hline 
LR  & 83.8 $\pm$0.09  & 97.51 $\pm$18.87  & 155.27 $\pm$45.56  & 54.17 $\pm$13.60  & 60.31 $\pm$21.17  & 0.01 $\pm$0.001\tabularnewline
\hline 
SVM  & 87.4 $\pm$0.07  & 78.74 $\pm$22.25  & 147.30 $\pm$57.99  & 42.43 $\pm$7.83  & 53.29 $\pm$9.35  & 0.06 $\pm$0.016\tabularnewline
\hline 
\end{tabular}
}
\label{Table:case3results}
\end{table*}

\begin{table*}
\caption{Predicting Total sales for PublishingCo-UDLU11 (medium number of different books sold)). Best values in bold.}
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE  & RMSE  & RAE (\%)  & RRSE (\%)  & Time (s)\tabularnewline
\hline 
\hline 
M5P  & 77.5 $\pm$0.15  & 573.41 $\pm$192.22  & 860.31 $\pm$456.22  & 58.26 $\pm$15.87  & 66.00 $\pm$29.45  & 0.024 $\pm$0.002\tabularnewline
\hline 
kNN  & 73.2 $\pm$0.16  & 596.62 $\pm$205.37  & 933.06 $\pm$356.11  & 60.62 $\pm$17.85  & 71.36 $\pm$22.34  &  \textbf{0 $\pm$0.0002}\tabularnewline
\hline 
RF &  \textbf{86.7 $\pm$0.07}  &  \textbf{449.16 $\pm$127.80}  &  \textbf{671.76 $\pm$212.74}  &  \textbf{45.58 $\pm$9.80}  &  \textbf{51.37 $\pm$11.16}  & 0.06 $\pm$0.01\tabularnewline
\hline 
LR  & 73.3 $\pm$0.20  & 680.16 $\pm$263.79  & 1103.97 $\pm$850.69  & 69.78 $\pm$27.24  & 86.11 $\pm$71.32  & 0.002 $\pm$0.001\tabularnewline
\hline 
SVM  & 80.4 $\pm$0.12  & 529.35 $\pm$172.50  & 846.01 $\pm$341.12  & 53.26 $\pm$11.92  & 63.54 $\pm$18.25  & 0.03 $\pm$0.02\tabularnewline
\hline 
\end{tabular}
}
\label{Table:case4results}
\end{table*}

%--------------------------------------------------------------------------

%\subsection{Sales forecasting with feature selection}
%\label{subsec:case_sales_feature_selection}

%***TO DO***: explain we have chosen the best Feature selection method (ReliefAttributeEval???). The rest of tables have been moved to an appendix.
%Sam will do it - JJ

In this stage of the experiments for the special cases, the
feature selection methods (RFE, CSE and SOM) are applied for each case
independently, since publishers with different sales ranges, and
targets, might in fact include different factors or have different
leverage when deciding how much to print. For clarity purpose, we present here only the results
after applying the RFE method for the four cases in Tables
\ref{Table:case1results-RAE}, \ref{Table:case2results-RAE},
\ref{Table:case3results-RAE} and \ref{Table:case4results-RAE}, this
one was chosen since the selected features by this
method showed in the previous experiment the highest predictive power among all classifiers. 
%According to the aforementioned tables, 
%It's not according to a table. It's according to the results there - JJ
RFE reduction has noticeably improved the performance of all classifiers specially in the case of the best selling publisher house (i.e. UDL929) and the
publisher which publishes a wide range of different books (i.e. UDL704). For the rest of the results which are obtained by CSE and SOM feature selection methods we refer the reader to the appendix of tables at the end of this paper.

\begin{table*}[!ht]
% Why the capital in Total? If it's a variable use Small Caps or Sans
% Serif - JJ
\caption{Predicting Total sales for PublishingCo-UDL929 (ReliefFAttributeEval-FS). Best values in bold.}
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE & RMSE & RAE (\%) & RRSE (\%) & Time (s)\tabularnewline
\hline 
\hline 
M5P &  \textbf{90.5 $\pm$0.1} &  \textbf{661.81 $\pm$186.17} &  \textbf{1070.24 $\pm$425.66} &  \textbf{40.18 $\pm$10.50} &  \textbf{43.56 $\pm$19.77} & 0.02 $\pm$0.002\tabularnewline
\hline 
kNN & 87.8 $\pm$0.11 & 751.54 $\pm$270.91 & 1257.82 $\pm$615.66& 44.57 $\pm$10.89 & 48.62 $\pm$18.73 &  \textbf{0.0001 $\pm$0.0003}\tabularnewline
\hline 
RF & 88.3 $\pm$0.11 & 735.04 $\pm$326.9 & 1420.43 $\pm$941.62 & 43.56 $\pm$15.88 & 54.36 $\pm$35.31 & 0.10 $\pm$0.009\tabularnewline
\hline 
LR & 88.4 $\pm$0.11 & 868.18 $\pm$208.65 & 1231.23 $\pm$454.86 & 53.38 $\pm$14.67 & 50.60 $\pm$22.04 & 0.0006 $\pm$0.0005\tabularnewline
\hline 
SVM & 87.6 $\pm$0.10 & 860.49 $\pm$334.67 & 1560.05 $\pm$962.64 & 49.98 $\pm$9.68 & 55.07 $\pm$17.76 & 0.023 $\pm$0.003\tabularnewline
\hline 
\end{tabular}
}
\label{Table:case1results-RAE}
\end{table*}

\begin{table*}[!ht]
\caption{Predicting Total sales for PublishingCo-UDLW12 (ReliefFAttributeEval-FS). Best values in bold.}
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE & RMSE & RAE (\%) & RRSE (\%) & Time (s)\tabularnewline
\hline 
\hline 
M5P & 83.7 $\pm$0.24 & 13.85 $\pm$6.56 & 23.19 $\pm$11.80 & 42.35 $\pm$20.33 & 54.08 $\pm$30.06 & 0.0063 $\pm$0.0014\tabularnewline
\hline 
kNN & 77.2 $\pm$0.30 & 14.18 $\pm$6.63 & 23.65 $\pm$11.03 & 43.37 $\pm$19.23 & 57.00 $\pm$30.59 &  \textbf{0 $\pm$0.0001}\tabularnewline
\hline 
RF &  \textbf{86.7 $\pm$0.16} & 13.63 $\pm$6.67 & 22.80 $\pm$11.81 & 41.07 $\pm$17.71 & 52.43 $\pm$26.23 & 0.0304 $\pm$0.0025\tabularnewline
\hline 
LR & 84.4 $\pm$0.25 &  \textbf{12.79 $\pm$6.65} &  \textbf{21.52 $\pm$12.54} & 39.47 $\pm$23.35 &  \textbf{50.14 $\pm$35.26} & 0.0005 $\pm$0.0005\tabularnewline
\hline 
SVM & 85.2 $\pm$0.19 & 13.22 $\pm$7.39 & 24.41 $\pm$15.09 &  \textbf{38.24 $\pm$15.41} & 50.46 $\pm$17.97 & 0.014 $\pm$0.002\tabularnewline
\hline 
\end{tabular}
}
\label{Table:case2results-RAE}
\end{table*}

\begin{table*}[!ht]
\caption{Predicting Total sales for PublishingCo704 (ReliefFAttributeEval-FS). Best values in bold. }
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE  & RMSE  & RAE (\%)  & RRSE (\%)  & Time (s)\tabularnewline
\hline 
\hline 
M5P  &  \textbf{89.3 $\pm$0.08}  &  \textbf{69.70 $\pm$19.85}  &  \textbf{122.48 $\pm$47.28}  & {38.49 $\pm$11.53}  &  \textbf{46.68 $\pm$16.02} & 0.028 $\pm$0.003\tabularnewline
\hline 
kNN  & 85.6 $\pm$0.10  & 78.39 $\pm$24.12  & 147.07 $\pm$58.02  & 42.58 $\pm$10.99  & 54.42 $\pm$15.54  &  \textbf{0.0001 $\pm$0.0003}\tabularnewline
\hline 
RF & 87.7 $\pm$0.08  & 73.13 $\pm$22.95  & 137.60 $\pm$56.14  & 39.79 $\pm$10.69  & 51.35 $\pm$17.00  & 0.13 $\pm$0.01\tabularnewline
\hline 
LR  & 87.3 $\pm$0.06  & 89.72 $\pm$18.31  & 140.65 $\pm$47.12  & 49.45 $\pm$10.50  & 53.20 $\pm$14.16  & 0.0008 $\pm$0.0005\tabularnewline
\hline 
SVM  & 87.7 $\pm$0.07  & 77.30 $\pm$22.25  & 146.46 $\pm$57.68  & 41.67 $\pm$7.99  & 53.10 $\pm$9.51  & 0.05 $\pm$0.02\tabularnewline
\hline 
\end{tabular}
}
\label{Table:case3results-RAE}
\end{table*}

\begin{table*}[!ht]
\caption{Predicting Total sales for PublishingCo-UDLU11 (ReliefFAttributeEval-FS). Best values in bold.}
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE & RMSE & RAE (\%) & RRSE (\%) & Time (s)\tabularnewline
\hline 
\hline 
M5P & 76.6 $\pm$0.17 & 592.49 $\pm$209.25 & 894.76 $\pm$508.14 & 60.21 $\pm$18.22 & 69.14 $\pm$37.35 & 0.02$\pm$0.001\tabularnewline
\hline 
kNN & 82.7 $\pm$0.09 & 487.00 $\pm$140.27 & 741.81 $\pm$238.56 & 49.27 $\pm$10.50 & 56.62 $\pm$12.59 &  \textbf{0.0001 $\pm$0.0003}\tabularnewline
\hline 
RF &  \textbf{85.3 $\pm$0.09} &  \textbf{455.7 $\pm$123.96} &  \textbf{686.54 $\pm$198.16} &  \textbf{46.32 $\pm$10.20} &  \textbf{53.03 $\pm$13.05} & 0.07 $\pm$0.003\tabularnewline
\hline 
LR & 64.4 $\pm$0.20 & 817.21 $\pm$331.16 & 1324.05 $\pm$1090.85 & 83.22 $\pm$32.01 & 101.90 $\pm$87.83 & 0.0005 $\pm$0.0005\tabularnewline
\hline 
SVM & 71.4 $\pm$0.14 & 644.42 $\pm$203.43 & 1009.66 $\pm$371.83 & 64.47 $\pm$11.70 & 75.54 $\pm$15.99 & 0.018 $\pm$0.001\tabularnewline
\hline 
\end{tabular}
}
\label{Table:case4results-RAE}
\end{table*}

%--------------------------------------------------------------------------

\subsection{Case study: publisher with the most sales}
\label{subsec:decision_trees}

What is, then, the model obtained with our method? What factors have
the biggest influence in sales and how can a publisher use them to gauge
sales of a new book? In order to do that, we have examined the models
built during the regression. Namely we present two of the decision
trees generated by M5P method applied to the data of the publisher
that sells the most. By looking at these trees, the publisher should
be able to decide how many actual books should be printed to put them
for sale, which was the initial objective of our research.

The model obtained for the publishing company UDL929 is shown in
Figure \ref{fig:decision_tree_best_seller}, which was the company that
sold the most books in our data set. It makes sense to use data for a
single company, since book sales for all companies is only available
if you buy it; a single company, however, will have complete data on
its own sales. This company is also representative because, being the
one that sells the most, will have the biggest range of sales for
individual books. In this figure, tree {\em leaves} correspond to
linear regression models that can be easily computed to obtain the
final predicted value for book sales.
The second decision tree (Figure
\ref{fig:decision_tree_best_seller_fs}) corresponds to the same
publishing company, so the same data were used, but the RAE feature
selection method was applied. This simplification, besides making the
tree easier to understand, reduces training time to a fraction of what
was needed before it.

Examining both trees, it can be noticed that
\texttt{tot\_points\_sale} variable appears twice in the trees, being
also the root. This indicates that the total number of points of sale
has the biggest influence on the decision models and also a major role in
forecasting the total sales, implying that buyers have to actually
{\em see}, maybe browse the book, before buying it. Books have to
physically {\em be} in the point of sale, so this number gives a lower
bound to the number of books that have to be printed; a stack of book
will have to be present in each point of sale to attain that effect. 
 This is an interesting clue for the
experts, who could check the expected impact of different number of
selling points. This feature also shows that there are different
regimes depending on the number of points of sale, with best sellers
being those that are distributed to more than roughly 500 points.

The second {\em splitting} feature, number of books distributed as a
novelty, gives the publisher a hint on how many books have to be sent
to bookstores to stack on the {\em novelty} table. You need to sell
roughly 5, the result of dividing 2807.5 y 550.5, to have a lower
bound in sales at 1006, which is negatively impacted by the retail
price and the number of points of sale in the first year of
distribution. In all cases, the last number in the formula featured in
the box gives you a baseline for sales, which the rest of the
variables impacting positively or negatively. A whole treatise could
be written on this decision tree but, in this case, let us look a bit
more closely at this box and try and interpret it. For instance, price
in this case will impact negatively on sales. For a typical retail
price of a trade paperback of 12 euro, we would be looking at negative
sales (if we disregard the rest of the variables) , which in this case
will indicate that these are the kind of books, maybe paperbacks, that
sell for a low price of around 5 euro. You also have to look at the
relation between the factor for total points of sale and the total
point of sales the first year. The fact that this last one is bigger
indicates that the publisher will have to {\em increase} the number of
points of sale after the first year, distributing it more shops at a
more capillary level; the number of points of sale will have to be
increased by 10\% to achieve a bit more than the baseline. All in all,
the conclusion with this part of the tree is that you will have to
print around 3000 books to sell a quantity that will start at 1000,
one third of the print run, and changes depending on a number of
factors, including price.

The right hand side of the tree include those books that are
distributed in a bigger percentage of the total amount of point of
sale available in Spain, which in Spain includes not only proper
bookstores, but also press stands and even supermarkets and hovers
around several tens of thousands (around 24000, according to
\cite{POS}), so we are clearly talking best sellers here distributed in
as many places as possible. The first tree
\ref{fig:decision_tree_best_seller} gives an idea of what we
are talking about, sales start at 1000 but the single factor that has
a bigger influence in sales is the topic, which has to be either 23 or
1. '1' is the first digit of the classification used for all kind of
general literary books. '23' includes philology, which includes either
textbooks of maybe books in a foreign language; there are just a few
of those, with '1' being far more common. So in principle these would
be the typical best-sellers, literary prizes and big hits, but we
would have to look more closely at the second tree
\ref{fig:decision_tree_best_seller_fs} to check what are other factors
influencing sales. It is also interesting that the number of total
points of sale comes multiplied by 7, which would be a good baseline
of books sold per outlet and give the publisher with a ballpark,
around 4000 books, to print in the first print run. The picture
painted by the second tree is a bit more bleak; obviously, even big
publishing house produce flops from time to time, and those would be
books which eventually print less than around 12500, indicating that
there has not been a second print run, since best-seller print runs
used to start at 15000 books\footnote{Nowadays in Spain that number
  would be closer to 5000}. In general, that branch of the tree paints
a pretty bleak picture: no matter how you combine the variable values,
it will end up with sales much lower than the print run. The main
problem with this tree is that its predictive value, with the
variables that we have, is not too high, except for the hint we have
on the subject matter. At the end of the day, there is a certain
amount of unpredictability in how much a book sells and the only
information we can give the publisher if he finds books sales adapt to
this model is to boost points of sale and lower price so that sales
pick a bit; in this sense and in certain cases our model is more
informative than predictive, but having numeric models helps anyway
the publisher with tailoring print runs and even increasing sales of
books that have been already printed.

The last box on the right of Figure
\ref{fig:decision_tree_best_seller_fs} has a negative baseline, but
this only indicates that the influence of the print run will be
bigger. This probably indicates books that have been reprinted due to
demand and it also tells how many are sold in every outlet: around 29
books. The price, number of initial points of sale, number of books
distributed as novelty and gifts all have a negative factor, but this
is related to the fact that they will have to be selling for a long
time, that is, not only be best, but also long sellers, which is also
indicated by the positive factor in the number of weeks they are for
sale.

All in all, these seven groups revealed by the tree in Figure
\ref{fig:decision_tree_best_seller_fs} constitute not only an accurate
model for predicting book sales based on data available to the
publisher, but also an informative tool that will allow publishers to
estimate print runs based on forecasted sales, and establish other
factors such as price, number of points of sale, and number of books
distributed to outlets.

An important fact that these tree models reveal is that marketing
variables, such as the total number of points of sale and the number
of units distributed as novelty have higher influence on the
prediction than the selling price of the book, which could be,
initially, considered as one of the most important parameters to
determine sales. This interpretation leads us to suggest that
companies could probably improve the accuracy of their sales
forecasting models by incorporating more marketing related variables
in their models. This does not mean that the retail price is not a
factor; it appears in all models, but it will be a secondary factor
that will have to be decided once the main ones are settled.

\begin{figure*}[!ht] 
\begin{center}
  \epsfig{file=decision_tree_no_fs.eps,width=13cm}
\caption{Decision tree built by the M5P method for the best seller publishing company including all pre-sales features.}
\label{fig:decision_tree_best_seller}
\end{center}
\end{figure*}

\begin{figure*}[!ht] 
\begin{center}
  \epsfig{file=decision_tree_fs.eps,width=13cm}
\caption{Decision tree built by the M5P method for the best seller
  publishing company using  the subset of features selected by the
  {\sf ReliefAttributeEval} method.} 
\label{fig:decision_tree_best_seller_fs}
\end{center}
\end{figure*}

%********************************************************************************
\section{Conclusions and Future Work}
\label{sec:conclusionsAndFutureWork}

In this paper, the problem of 
predicting total sales in order to print the right amount of books
%how many books should be printed 
when a new book is going to be published is faced using several data-mining and forecasting methods.

This is a very challenging problem in this industry, because printing
a much higher 
number of volumes than those finally sold will lead to losses, while
printing an adequate number of copies will optimise sales and company
earnings.  
In addition, there are several difficulties inherent to the new book
sales forecasting, such as the limited amount of historical data to
consider or the variability of the market (fashions, seasonality),
which do a difficult task finding an adequate forecasting method.

A dataset consisting of 6000 books provided by the Spanish publishing
company Trevenque Editorial S.L. has been used. This data has been
analysed by means of Self-Organizing Maps as a visualisation tool, and
two other Feature Selection techniques, namely a correlation-based
feature subset evaluation method, and an adaptation of relief for
attribute estimation. Using these methods the most relevant variables
when predicting sales have been determined, while unrelated or
superfluous ones have been removed.

Four different datasets have also been extracted from the original
one, each one corresponding to a publishing house which we have
considered as representative: the one with the most number of sales,
the one with the medium number of sales, with the most varied sales
(different books), and with a medium number of different books sold.

Considering these datasets and the complete one, five machine learning
models for sales forecasting have been configured (according to the
problem requirements) and used to predict new book sales, from which
the optimal print-run could be estimated. These algorithms include
some based on Decision Trees (M5P and Random Forest), kNN, Support
Vector Machines and Linear Regression. 
These methods have been evaluated considering several performance
indicators (error measures) and also the time consumed to build the
corresponding models.

As a conclusion, M5P yields the best results overall, obtaining the best accuracy 
(lowest error) on average, even with acceptable values regarding the training time.
In addition, the Decision Trees obtained with this method have been
shown to be an excellent decision-aid tool for publishers, since they are accurate 
and quite easy to understand and apply. 
They are also able to provide insights on the decision variables that
have a higher influence on the predicted sales for these publishers.

This, compared to the output of black-box models such as artificial
neural networks, was a determining advantage from the point of view of
the needs of experts and the company that initially contacted us for
this work.

The feature selection techniques have also shown to positively affect
all the results, yielding a better performance in time to train while
maintaining good accuracy. Moreover, their application leads to obtain
even simpler models.

Overall, this study contributes to the literature by proposing the use
of different soft-computing standard techniques to solve a new
challenging problem. 
Moreover, proposed method, consisting in selecting features and
applying different models for different types of publishers could be
implemented, not only in the publishing  
industry, but also in other domains where the specificity of products
is similar such as novelty items or toys, and might be of interest to other academic researches and
industrial practitioners.  
% Antonio - I don't know if we should say this...

%Thus, the methodology presented here provides a reliable data processing technique for editorial data and also yields a reasonable result that allows publishers to forecast sales and then create the appropriate print run without incurring in losses.

% -------- Future Work ---------

As future work, it would be interesting to compare the obtained
forecasting results with those yielded by other soft-computing
methods, such as genetic programming or fuzzy logic. 
Moreover, even if the methods have been configured to obtain the best
forecasting results, it would be of interest carrying out some
automatic parameter-tuning procedure by means of optimisation
meta-algorithms \cite{huang2006ga,Mookiah2013_EA_Tuning,Shen2016_FFly_Tuning},  
% Antonio - changed to include KBS references
for instance.

In addition, if data for other countries were available, it would be
interesting to observe, and factor in the method, country differences,
even considering the same book and how it sells in different
countries.

Furthermore, since there are some characteristics that may affect product sales 
significantly, another way to improve sales predictions could be making a 
thorough study of the effect of the application of discounts and promotions.

Finally, there are some variables over which the publisher has some
control once the book has been published, such as the initial print-run, 
the books offered as gifts to reviewers, or what proportion is going to be sent 
to each possible point of sale. These can be incorporated into an optimisation 
model that will optimise those values to obtain desired sales.

%********************************************************************************
\section*{Acknowledgements}

This work has been supported in part by projects PreTEL (PRM Consultores - Trevenque S.L.), TIN2014-56494-C4-3-P and TEC2015-68752 (Spanish Ministry of Economy and Competitiveness and FEDER), PRY142/14 (Fundaci{\'o}n P{\'u}blica Andaluza Centro de Estudios Andaluces en la IX Convocatoria de Proyectos de Investigaci{\'o}n), and PROY-PP2015-06 (Plan Propio 2015, funded by the University of Granada, Spain).

%********************************************************************************
\bibliographystyle{elsarticle-num}
%\bibliography{refs}
\begin{thebibliography}{10}
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\expandafter\ifx\csname href\endcsname\relax
  \def\href#1#2{#2} \def\path#1{#1}\fi

\bibitem{Fildes2010}
R.~Fildes, B.~Kingsman, {Incorporating demand uncertainty and forecast error in
  supply chain planning models}, J Oper Res Soc 62 (2010) 483--500.

\bibitem{Saeed2008}
K.~Saeed, {Trend forecasting for stability in supply chains}, J Bus Res 61
  (2008) 1113--1124.

\bibitem{Zhao2001}
X.~Zhao, J.~Xie, R.~Lau, {Improving the supply chain performance: use of
  forecasting models versus early order commitments}, Int J Prod Res 39~(17)
  (2001) 3923--3939.

\bibitem{Little1998}
J.~Little, {Integrated measures of sales, merchandising and distribution}, Int
  J Res Mark 15 (1998) 473--485.

\bibitem{Armstrong2001}
J.~Armstrong, {Principles of forecasting. A handbook for researchers and
  practitioners}, Kluwer, Norwell, 2001.

\bibitem{SThomassey2014}
S.~Thomassey, {Sales Forecasting in Apparel and Fashion Industry: A Review},
  Intell Fashion Forecasting Syst: Models and App (2014) 2--27.

\bibitem{DeToni2000}
A.~D. Toni, A.~Meneghetti, {The production planning process for a network of
  firms in the textile-apparel industry}, Int J Prod Econ 65 (2000) 17--32.

\bibitem{kohonen1998}
T.~Kohonen, {The self-organizing map}, Neurocomputing 21~(1-3) (1998) 1--6.

\bibitem{Chang2009}
P.~Chang, C.~Liu, C.~Fan, {Data clustering and fuzzy neural network for sales
  forecasting: A case study in printed circuit board industry}, Knowl Based
  Syst (2009) 344--355.

\bibitem{Sanders1994}
N.~Sanders, K.~Manrodt, {Forecasting practices in US corporations: Survey
  results}, Interfaces 24~(2) (1994) 92--100.

\bibitem{Papalexopoulos1990}
A.~Papalexopoulos, T.~Hesterberg, {A regression-based approach to short-term
  system load forecasting}, IEEE Trans Power Syst 5 (1990) 1535--1547.

\bibitem{Yoo1999}
H.~Yoo, R.~Pimmel, {Short-term load forecasting using a self-supervised
  adaptive neural network}, IEEE Trans Power Syst 14~(2) (1999) 779--784.

\bibitem{Mastorocostas2001}
P.~Mastorocostas, J.~Theocharis, V.~Petridis, {A constrained orthogonal
  least-squares method for generating TSK fuzzy models: application to
  short-term load forecasting}, Fuzzy Set Syst 118 (2001) 215--233.

\bibitem{Chu2003}
C.~Chu, P.~Zhang, {A comparative study of linear and nonlinear models for
  aggregate retail sales forecasting}, Int J Prod Econ 86 (2003) 217--231.

\bibitem{Brown1959}
R.~Brown, {Smoothing forecasting and prediction of discrete time series},
  Prentice Hall, Englewood Cliffs, 1959.

\bibitem{Winters1960}
P.~Winters, {Forecasting sales by exponential weighed moving averages}, Manag
  Sci 6 (1960) 324--342.

\bibitem{Box1969}
G.~Box, G.~Jenkins, {Time series analysis forecasting and control}, Prentice
  Hall, Englewood Cliffs, 1969.

\bibitem{KayacanUK10}
E.~Kayacan, B.~Ulutas, O.~Kaynak, Grey system theory-based models in time
  series prediction, Expert Syst Appl 37~(2) (2010) 1784--1789.

\bibitem{ChingChin2010}
C.~Ching-Chin, A.~I.~K. Ieng, W.~Ling-Ling, K.~Ling-Chieh, {Designing a
  decision-support system for new product sales forecasting}, Expert Syst Appl
  37~(2) (2010) 1654--1665.

\bibitem{FaderHardie2005}
S.~Fader, B.~Hardie, {The value of simple models in new product forecasting and
  customer-base analysis}, Appl Stoch Model Bus (2005) 461--473.

\bibitem{Madsen2008}
H.~Madsen, {Time series analysis}, Chapman and Hall. CRC Co., 2008.

\bibitem{Lapide1999}
L.~Lapide, {New developments in business forecasting}, J Bus Forecasting (1999)
  29--30.

\bibitem{ChernWSF15}
C.-C. Chern, C.-P. Wei, F.-Y. Shen, Y.-N. Fan, A sales forecasting model for
  consumer products based on the influence of online word-of-mouth., Inf Syst E
  Bus Manag 13 (2015) 445--473.

\bibitem{Moon2010ICSSSM}
G.~Moon, G.~Kikuta, T.~Yamada, A.~Yoshikawa, T.~Terano, {Blog information
  considered useful for book sales prediction}, in: 7th International
  Conference on Service Systems and Service Management (ICSSSM), IEEE. Tokyo,
  2010, pp. 1--5.

\bibitem{Moon2010ICEC}
G.~Moon, G.~Kikuta, T.~Yamada, A.~Yoshikawa, T.~Terano, Categorization of sales
  patterns to use blog information as book sales prediction, in: Proceedings of
  the 12th International Conference on Electronic Commerce: Roadmap for the
  Future of Electronic Business, ICEC '10, ACM, New York, NY, USA, 2010, pp.
  1--7.

\bibitem{Hammond1990}
J.~Hammond, {Quick response in apparel industries}, Havard Business school,
  Cambridge, 1990.

\bibitem{Xia2012}
M.~Xia, Y.~Zhang, L.~W.~S. Ye, {Fashion retailing forecasting based on extreme
  learning machine with adaptive metrics of inputs}, Knowl Based Syst 36 (2012)
  253--259.

\bibitem{Thomassey2012}
S.~Thomassey, {Sales forecasts in clothing industry. The key success factor of
  the supply chain management}, Int J Prod Econ 128 (2012) 470--483.

\bibitem{wiki:dewey}
Wikipedia,
  \href{\url{https://en.wikipedia.org/wiki/List_of_Dewey_Decimal_classes}}{List
  of dewey decimal classes}, [Online; accessed 05-May-2016].
\newline\urlprefix\url{\url{https://en.wikipedia.org/wiki/List_of_Dewey_Decimal_classes}}

\bibitem{fayyad1996data}
U.~Fayyad, G.~Piatetsky-Shapiro, P.~Smyth, From data mining to knowledge
  discovery in databases, AI magazine 17~(3) (1996) 37.

\bibitem{kittler1986feature}
J.~Kittler, Handbook of pattern recognition and image processing, Academic
  press, 1986.

\bibitem{Krishnapuram2004}
B.~Krishnapuram, A.~Hartemink, L.~Carin, M.~Figueiredo, {A Bayesian Approach to
  Joint Feature Selection and Classifier Design}, IEEE T Pattern Anal 26~(9)
  (2004) 1105--1111.

\bibitem{Chen2011_FS_PTS}
Y.~Chen, D.~Miao, R.~Wang, K.~Wu, A rough set approach to feature selection
  based on power set tree, Knowl Based Syst 24~(2) (2011) 275 -- 281.

\bibitem{Selvakuberan2008}
K.~Selvakuberan, M.~Indradevi, R.~Rajaram, {Combined Feature Selection and
  classification - A novel approach for the categorization of web pages}, J Inf
  Comp Sci 3~(2) (2008) 083--089.

\bibitem{liu1998feature}
H.~Liu, H.~Motoda, {Feature extraction, construction and selection: A data
  mining perspective}, Springer Science \& Business Media, 1998.

\bibitem{Hall2009}
M.~Hall, E.~Frank, G.~Holmes, B.~Pfahringer, P.~Reutemann, I.~Witten, {The WEKA
  data mining software: an update}, ACM SIGKDD Explor Newslett 11~(1) (2009)
  10--18.

\bibitem{Witten2011}
I.~Witten, E.~Frank, M.~Hall, {Data mining: Practical machine learning tools
  and techniques}, Morgan Kaufmann, 2011.

\bibitem{Hall1998}
M.~A. Hall, Correlation-based feature subset selection for machine learning,
  Ph.D. thesis, University of Waikato, Hamilton, New Zealand (1998).

\bibitem{RobnikSikonja1997}
M.~Robnik-Sikonja, I.~Kononenko, An adaptation of relief for attribute
  estimation in regression, in: D.~H. Fisher (Ed.), Fourteenth International
  Conference on Machine Learning, Morgan Kaufmann, 1997, pp. 296--304.

\bibitem{Kira1992}
K.~Kira, L.~A. Rendell, {A Practical Approach to Feature Selection}, in: D.~H.
  Sleeman, P.~Edwards (Eds.), Ninth International Workshop on Machine Learning,
  Morgan Kaufmann, 1992, pp. 249--256.

\bibitem{UmatUlts}
S.~Ultsch, Kohonen's self-organizing maps for exploratory data analysis, in:
  Proc. of the INNC'90, 2000, pp. 305--308.

\bibitem{Quinlan1986}
J.~Quinlan, {Induction of decision trees}, Mach Learn 1 (1986) 181--186.

\bibitem{Quinlan1992}
R.~J. Quinlan, Learning with continuous classes, in: 5th Australian Joint
  Conference on Artificial Intelligence, World Scientific, Singapore, 1992, pp.
  343--348.

\bibitem{Wang1997}
Y.~Wang, I.~H. Witten,
  \href{https://www.researchgate.net/profile/Ian_Witten/publication/33051395_Induction_of_model_trees_for_predicting_continuous_classes/links/545948ea0cf2cf516483ce67.pdf}{Induction
  of model trees for predicting continuous classes}, Tech. rep., Dept. of
  Computer Science, University of Waikato, working Paper 96/23, Poster papers
  of the 9th European Conference on Machine Learning (1996).
\newline\urlprefix\url{https://www.researchgate.net/profile/Ian_Witten/publication/33051395_Induction_of_model_trees_for_predicting_continuous_classes/links/545948ea0cf2cf516483ce67.pdf}

\bibitem{WittenFrank2000}
I.~Witten, E.~Frank, {Data mining}, Morgan Kaufmann, 2000.

\bibitem{Solomatine2004}
D.~Solomatine, Y.~Xue, {M5 Model Trees and Neural Networks: Application to
  FloodForecasting in the Upper Reach of the Huai River in China}, J Hydrol Eng
  9~(6) (2004) 491--501.

\bibitem{Aha1991}
D.~Aha, D.~Kibler, Instance-based learning algorithms, Mach Learn 6 (1991)
  37--66.

\bibitem{Mitchell1997}
T.~Mitchell, {Machine Learning}, McGraw-Hill Companies, Inc., 1997.

\bibitem{Breiman2001}
L.~Breiman, Random forests, Mach Learn 45~(1) (2001) 5--32.

\bibitem{Cohen2003}
J.~Cohen, P.~Cohen, S.~West, L.~Aiken, {Applied multiple regression/correlation
  analysis for the behavioral sciences}, Lawrence Erlbaum Associates, 2003.

\bibitem{Yan2009}
X.~Yan, {Linear Regression Analysis: Theory and Computing}, World Scientific,
  2009.

\bibitem{Rencher2102}
A.~Rencher, W.~Christensen, {Multivariate regression}, Vol. 709 of Wiley Series
  in Probability and Statistics, John Wiley and Sons, 2012.

\bibitem{EnkeT05}
D.~Enke, S.~Thawornwong, The use of data mining and neural networks for
  forecasting stock market returns, Expert Syst Appl 29~(4) (2005) 927--940.

\bibitem{McClendon2015}
L.~McClendon, N.~Meghanathan, Using machine learning algorithms to analyze
  crime data, Mach Learn and App 2~(1).

\bibitem{Cortes1995}
C.~Cortes, V.~Vapnik, {Support Vector Networks}, Mach Learn 20 (1995) 273--297.

\bibitem{Shevade1999}
S.~Shevade, S.~Keerthi, C.~Bhattacharyya, K.~Murthy, Improvements to the smo
  algorithm for svm regression, IEEE T Neural Networ.

\bibitem{MinSVM05}
J.~H. Min, Y.~Lee, Bankruptcy prediction using support vector machine with
  optimal choice of kernel function parameters, Expert Syst. Appl. 28~(4)
  (2005) 603--614.

\bibitem{Cao2003}
L.~Cao, {Support vector machine experts for time series forcasting},
  Neurocomputing 51 (2003) 321--339.

\bibitem{Jari2008}
C.~Jari, L.~Wen, L.~Kang, {Support vector machine classification for large data
  set vis minmum enclosing ball clustering}, Neurocomputing 71 (2008) 611--619.

\bibitem{gepsoft}
{Gepsoft}, {Analyzing GeneXproTools Models Statistically. Statistical Analysis
  of Time Series Prediction Models}, {\tt
  http://www.gepsoft.com/gxpt4kb/Chapter10/Section3/Introduction.htm} (Dec.
  2015).

\bibitem{pentaho}
{pentaho}, {Pentaho Data Mining Community Documentation}, {\tt
  http://wiki.pentaho.com/display/DATAMINING} (Dec. 2015).

\bibitem{otexts}
{otexts}, {Forecasting: principles and practice. The forecaster's toolbox.
  Evaluating forecast accuracy}, {\tt https://www.otexts.org/fpp/2/5} (Dec.
  2015).

\bibitem{Hyndman2013}
R.~Hyndman, G.~Athanasopoulos, Forecasting: principles and practice, OTexts,
  2013.

\bibitem{yu2004ec}
X.~Yu, S.~Liong, V.~Babovic, Ec-svm approach for real-time hydrologic
  forecasting., J Hydroinform 6 (2004) 209--223.

\bibitem{chao2015construction}
C.-F. Chao, M.-H. Horng, The construction of support vector machine classifier
  using the firefly algorithm, Computat Intell Neurosci.

\bibitem{hornik1989multilayer}
K.~Hornik, M.~Stinchcombe, H.~White, Multilayer feedforward networks are
  universal approximators, Neural networks 2~(5) (1989) 359--366.

\bibitem{mirjalili2014let}
S.~Mirjalili, S.~M. Mirjalili, A.~Lewis, Let a biogeography-based optimizer
  train your multi-layer perceptron, Inform Sciences 269 (2014) 188--209.

\bibitem{wdaa2008differential}
A.~S.~I. Wdaa, Differential evolution for neural networks learning enhancement,
  Ph.D. thesis, Universiti Teknologi Malaysia (2008).

\bibitem{mirjalili2015effective}
S.~Mirjalili, \href{http://dx.doi.org/10.1007/s10489-014-0645-7}{How effective
  is the grey wolf optimizer in training multi-layer perceptrons}, Appl Intell
  43~(1) (2015) 150--161.
\newblock \href {http://dx.doi.org/10.1007/s10489-014-0645-7}
  {\path{doi:10.1007/s10489-014-0645-7}}.
\newline\urlprefix\url{http://dx.doi.org/10.1007/s10489-014-0645-7}

\bibitem{Geisser1993}
S.~Geisser, {Predictive Inference}, Chapman and Hall, 1993.

\bibitem{Kohavi1995}
R.~Kohavi, A study of cross-validation and bootstrap for accuracy estimation
  and model selection, in: Proceedings of the Fourteenth International Joint
  Conference on Artificial Intelligence, {IJCAI} 95, Montr{\'{e}}al
  Qu{\'{e}}bec, Canada, August 20-25 1995, 2 Volumes, Morgan Kaufmann, 1995,
  pp. 1137--1145.

\bibitem{Devijver1982}
P.~Devijver, J.~Kittler, {Pattern Recognition: A Statistical Approach},
  Prentice-Hall, 1982.

\bibitem{POS}
M.~Gil, Puntos de venta de prensa y libros,
  \url{https://antinomiaslibro.wordpress.com/2013/11/21/puntos-de-venta/},
  consulted on May 21st, 2016 (November 2013).

\bibitem{huang2006ga}
C.-L. Huang, C.-J. Wang, A ga-based feature selection and parameters
  optimization for support vector machines, Expert Syst Appl 31~(2) (2006)
  231--240.

\bibitem{Mookiah2013_EA_Tuning}
M.~Mookiah, U.~Acharya, R.~Martis, C.~Chua, C.~Lim, E.~Ng, A.~Laude,
  Evolutionary algorithm based classifier parameter tuning for automatic
  diabetic retinopathy grading: A hybrid feature extraction approach, Knowl
  Based Syst 39 (2013) 9--22.

\bibitem{Shen2016_FFly_Tuning}
L.~Shen, H.~Chen, Z.~Yu, W.~Kang, B.~Zhang, H.~Li, B.~Yang, D.~Liu, Evolving
  support vector machines using fruit fly optimization for medical data
  classification, Knowledge-Based Systems 96 (2016) 61--75.

\end{thebibliography}

%----------------------------------------------------------------------

\appendix
\section{APENDIX TABLES}

\begin{table*}[!ht]
\caption{Predicting Total sales for PublishingCo704 (CfsSubsetEval-FS). Best values in bold.}
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE & RMSE & RAE (\%) & RRSE (\%) & Time (s)\tabularnewline
\hline 
\hline 
M5P & 80.1 $\pm$0.09 & 104.87 $\pm$21.38 & 174.57 $\pm$47.33 & 58.11 $\pm$14.73 & 67.58 $\pm$22.61 & 0.018 $\pm$0.003\tabularnewline
\hline 
kNN & 78.2 $\pm$0.12 & 99.45 $\pm$24.31 & \textbf{170.83 $\pm$46.34} & 54.53 $\pm$12.28 & 65.86 $\pm$18.34 & \textbf{0.0001 $\pm$0.0003}\tabularnewline
\hline 
RF & 77.6 $\pm$0.13 & 101.93 $\pm$23.66 & 180.56 $\pm$51.20 & 56.10 $\pm$13.62 & 69.86 $\pm$23.27 & 0.0738 $\pm$0.003\tabularnewline
\hline 
LR & 80.1 $\pm$0.10 & 106.62 $\pm$21.52 & 173.18 $\pm$46.58 & 59.11 $\pm$14.96 & 67.53 $\pm$24.53 & 0.0004 $\pm$0.0005\tabularnewline
\hline 
SVM & \textbf{81.8 $\pm$0.08} & \textbf{94.29 $\pm$25.52} & 180.54 $\pm$62.67 & \textbf{50.79 $\pm$8.48} & \textbf{65.62 $\pm$6.25} & 0.040 $\pm$0.02\tabularnewline
\hline 
\end{tabular}
}
\end{table*}

\begin{table*}[!ht]
\caption{Predicting Total sales for PublishingCo704 (SOM-FS). Best values in bold.}
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE & RMSE & RAE (\%) & RRSE (\%) & Time (s)\tabularnewline
\hline 
\hline 
M5P & \textbf{89.2 $\pm$0.07} & \textbf{71.536 $\pm$18.95} & \textbf{123.06 $\pm$42.13} & \textbf{39.43 $\pm$10.58} & \textbf{47.20 $\pm$15.16} & 0.031 $\pm$0.007\tabularnewline
\hline 
kNN & 85.8 $\pm$0.09 & 79.42 $\pm$21.47 & 140.95 $\pm$44.93 & 43.34 $\pm$10.07 & 53.31 $\pm$14.37 & \textbf{0.0001 $\pm$0.0002}\tabularnewline
\hline 
RF & 87.7 $\pm$0.08 & 73.98 $\pm$20.91 & 134.11 $\pm$51.70 & 40.48 $\pm$10.51 & 50.41 $\pm$16.38 & 0.10 $\pm$0.0076\tabularnewline
\hline 
LR & 87.5 $\pm$0.06 & 88.70 $\pm$16.34 & 135.03 $\pm$35.18 & 49.08 $\pm$10.47 & 51.97 $\pm$13.65 & 0.0006 $\pm$0.0005\tabularnewline
\hline 
SVM & 87.8 $\pm$0.06 & 81.09 $\pm$22.48 & 154.82 $\pm$59.93 & 43.65 $\pm$7.36 & 55.72 $\pm$7.15 & 0.031 $\pm$0.003\tabularnewline
\hline 
\end{tabular}
}
\end{table*}

%-----------------------------------

\begin{table*}[!ht]
\caption{Predicting Total sales for PublishingCo-UDLU11 (CfsSubsetEval-FS). Best values in bold.}
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE & RMSE & RAE (\%) & RRSE (\%) & Time (s)\tabularnewline
\hline 
\hline 
M5P & 64.0 $\pm$0.23 & 777.94 $\pm$342.04 & 1305.28 $\pm$1075.29 & 78.37 $\pm$30.4 & 99.86 $\pm$85.43 & 0.013 $\pm$0.001\tabularnewline
\hline 
kNN & 64.8 $\pm$0.18 & 702.70 $\pm$181.41 & \textbf{1008.39 $\pm$305.87} & 71.41 $\pm$13.95 & \textbf{78.47 $\pm$22.70} & \textbf{0.0001 $\pm$0.0002}\tabularnewline
\hline 
RF & 65.3 $\pm$0.19 & \textbf{685.56 $\pm$195.28} & 1021.37 $\pm$352.88 & \textbf{69.74 $\pm$16.96} & 79.37 $\pm$26.98 & 0.0466 $\pm$0.0023\tabularnewline
\hline 
LR & \textbf{67.0 $\pm$0.23}\ & 857.22 $\pm$308.69 & 1352.26 $\pm$945.31 & 86.10 $\pm$24.25 & 102.89 $\pm$74.43 & 0.0003 $\pm$0.0005\tabularnewline
\hline 
SVM & 64.8 $\pm$0.23 & 879.91 $\pm$280.33 & 1402.82 $\pm$593.44 & 87.74 $\pm$15.27 & 105.40 $\pm$37.73 & 0.017 $\pm$0.001\tabularnewline
\hline 
\end{tabular}
}
\end{table*}

\begin{table*}[!ht]
\caption{Predicting Total sales for PublishingCo-UDLU11 (SOM-FS). Best values in bold.}
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE & RMSE & RAE (\%) & RRSE (\%) & Time (s)\tabularnewline
\hline 
\hline 
M5P & 79.0 $\pm$0.14 & 573.25 $\pm$167.88 & 827.55 $\pm$388.16 & 58.42 $\pm$14.14 & 63.98 $\pm$25.63 & 0.0153 $\pm$0.0015\tabularnewline
\hline 
kNN & 83.6 $\pm$0.10 & 492.18 $\pm$126.86 & 727.33 $\pm$209.88 & 50.12 $\pm$10.69 & 56.14 $\pm$12.89 & \textbf{0 $\pm$0.0002}\tabularnewline
\hline 
RF & \textbf{85.6 $\pm$0.08} & \textbf{446.25 $\pm$119.79} & \textbf{673.46 $\pm$189.06} & \textbf{45.38 $\pm$9.78} & \textbf{52.24 $\pm$13.09} & 0.0594 $\pm$0.0049\tabularnewline
\hline 
LR & 71.0 $\pm$0.19 & 723.74 $\pm$247.96 & 1108.74 $\pm$736.17 & 74.14 $\pm$24.91 & 85.96 $\pm$59.99 & 0.0004 $\pm$0.0005\tabularnewline
\hline 
SVM & 71.6 $\pm$0.15 & 640.14 $\pm$206.46 & 1017.02 $\pm$376.13 & 63.96 $\pm$11.87 & 76.10 $\pm$16.80 & 0.019 $\pm$0.001\tabularnewline
\hline 
\end{tabular}
}
\end{table*}

%-----------------------------------

\begin{table*}[!ht]
\caption{Predicting Total sales for PublishingCo-UDL929 (CfsSubsetEval-FS). Best values in bold.}
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE & RMSE & RAE (\%) & RRSE (\%) & Time (s)\tabularnewline
\hline 
\hline 
M5P & \textbf{88.0 $\pm$0.13} & \textbf{738.28 $\pm$188.53} & \textbf{1141.79 $\pm$472.025} & \textbf{45.45 $\pm$13.62} & \textbf{47.58 $\pm$24.84} & 0.017 $\pm$0.002\tabularnewline
\hline 
kNN & 85.3 $\pm$0.15& 825.19 $\pm$235.34 & 1280.33 $\pm$523.06 & 50.23 $\pm$14.07 & 51.42 $\pm$21.24 & \textbf{0.0001 $\pm$0.0002}\tabularnewline
\hline 
RF & 86.4 $\pm$0.14 & 790.09 $\pm$225.86 & 1245.23 $\pm$584.23 & 48.22 $\pm$13.79 & 50.01 $\pm$21.97 & 0.063 $\pm$0.0039\tabularnewline
\hline 
LR & 86.6 $\pm$0.12 & 936.56 $\pm$204.46 & 1329.24 $\pm$480.79 & 58.04 $\pm$16.92 & 55.41 $\pm$25.62 & 0.0004 $\pm$0.0005\tabularnewline
\hline 
SVM & 86.8 $\pm$0.12 & 940.39 $\pm$344.16 & 1648.50 $\pm$1026.04 & 54.50 $\pm$7.12 & 56.17 $\pm$7.8 & 0.020 $\pm$0.002\tabularnewline
\hline 
\end{tabular}
}
\end{table*}

\begin{table*}[!ht]
\caption{Predicting Total sales for PublishingCo-UDL929 (SOM-FS). Best values in bold.}
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE & RMSE & RAE (\%) & RRSE (\%) & Time (s)\tabularnewline
\hline 
\hline 
M5P & \textbf{89.7 $\pm$0.10} & \textbf{685.67 $\pm$185.99} & \textbf{1116.42 $\pm$478.06} & \textbf{42.09 $\pm$12.8} & 46.31 $\pm$24.61 & 0.02 $\pm$0.004\tabularnewline
\hline 
kNN & 89.0 $\pm$0.10 & 746.45 $\pm$250.76 & 1203.41 $\pm$564.84 & 44.36 $\pm$9.73 & \textbf{46.25 $\pm$14.41} & \textbf{0.0001 $\pm$0.0003}\tabularnewline
\hline 
RF & 89.5 $\pm$0.09 & 708.24 $\pm$264.52 & 1270.26 $\pm$708.82 & 42.14 $\pm$12.45 & 49.07 $\pm$26.59 & 0.080 $\pm$0.0058\tabularnewline
\hline 
LR & 86.6 $\pm$0.13 & 894.08 $\pm$206.60 & 1282.83 $\pm$451.87 & 55.29 $\pm$15.92 & 53.74 $\pm$24.70 & 0.0005 $\pm$0.0005\tabularnewline
\hline 
SVM & 87.1 $\pm$0.10 & 908.56 $\pm$348.20 & 1608.77 $\pm$972.55 & 52.63 $\pm$8.75 & 55.95 $\pm$12.49 & 0.022 $\pm$0.002\tabularnewline
\hline 
\end{tabular}
}
\end{table*}

%-----------------------------------

\begin{table*}[!ht]
\caption{Predicting Total sales for PublishingCo-UDLW12 (CfsSubsetEval-FS). Best values in bold.}
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE & RMSE & RAE (\%) & RRSE (\%) & Time (s)\tabularnewline
\hline 
\hline 
M5P & 81.6 $\pm$0.22 & 15.67 $\pm$6.58 & 25.63 $\pm$11.84 & 47.93 $\pm$18.81 & 60.72 $\pm$31.75 & 0.006 $\pm$0.002\tabularnewline
\hline 
kNN & 75.2 $\pm$0.27 & 15.73 $\pm$7.73 & 26.68 $\pm$14.82 & 48.58 $\pm$24.77 & 66.23 $\pm$48.59 & \textbf{0.0001 $\pm$0.0003}\tabularnewline
\hline 
RF & 75.1 $\pm$0.26 & \textbf{15.61 $\pm$7.19} & 26.83 $\pm$13.36 & \textbf{47.38 $\pm$20.10} & 64.65 $\pm$38.66 & 0.0224 $\pm$0.0027\tabularnewline
\hline 
LR & 81.6 $\pm$0.22 & 15.67 $\pm$6.55 & \textbf{25.49 $\pm$11.63} & 47.97 $\pm$18.85 & \textbf{60.28 $\pm$30.62} & 0.0002 $\pm$0.0004\tabularnewline
\hline 
SVM & \textbf{81.8 $\pm$0.21} & 17.71 $\pm$9.70 & 31.95 $\pm$19.26 & 50.07 $\pm$14.93 & 65.12 $\pm$16.36 & 0.013 $\pm$0.001\tabularnewline
\hline 
\end{tabular}
}
\end{table*}

\begin{table*}[!ht]
\caption{Predicting Total sales for PublishingCo-UDLW12 (SOM-FS). Best values in bold.}
\centering{}%
\resizebox{12cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
 & $R^{2}$  & MAE & RMSE & RAE (\%) & RRSE (\%) & Time (s)\tabularnewline
\hline 
\hline 
M5P & \textbf{86.0 $\pm$0.19} & 13.84 $\pm$6.75 & 23.94 $\pm$11.64 & 41.65 $\pm$17.56 & 55.44 $\pm$27.13 & 0.006 $\pm$0.001\tabularnewline
\hline 
kNN & 74.0 $\pm$0.29 & 15.93 $\pm$8.52 & 28.28 $\pm$16.28 & 48.83 $\pm$25.89 & 68.75 $\pm$48.85 & \textbf{0 $\pm$0.0002}\tabularnewline
\hline 
RF & 85.5 $\pm$0.19 & \textbf{13.66 $\pm$6.47} & \textbf{22.51 $\pm$11.03} & \textbf{41.38 $\pm$17.87} & \textbf{52.20 $\pm$27.09} & 0.026 $\pm$0.0019\tabularnewline
\hline 
LR & 85.4 $\pm$0.21 & 14.59 $\pm$7.13 & 24.40 $\pm$12.87 & 44.81 $\pm$23.65 & 57.73 $\pm$38.63 & 0.0003 $\pm$0.0005\tabularnewline
\hline 
SVM & 84.4 $\pm$0.18 & 14.18 $\pm$7.96 & 26.49 $\pm$16.54 & 40.93 $\pm$16.65 & 54.37 $\pm$20.00 & 0.01 $\pm$0.001\tabularnewline
\hline 
\end{tabular}
}
\end{table*}

\end{document}
%%% Local Variables:
%%% ispell-local-dictionary: "english"
%%% End: